### 2019-01-26 paper.tex
Initial draft

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Large-scale peer-to-peer network for mobile platforms: Challenges and Experiences\\
\thanks{DAIS-ITA}
}

\author{\IEEEauthorblockN{Nirmit Desai}
\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
%\textit{}\\
Yorktown Heights, NY, USA \\
nirmit.desai@us.ibm.com}
\and
\IEEEauthorblockN{Wendy Chong}
\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
%\textit{name of organization (of Aff.)}\\
Yorktown Heights, NY, USA \\
wendych@us.ibm.com}
\and
\IEEEauthorblockN{Heather Achilles}
\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
%\textit{name of organization (of Af}\\
Yorktown Heights, NY, USA \\
hachilles@us.ibm.com}
\and
\IEEEauthorblockN{Thomas La Porta}
\IEEEauthorblockA{\textit{Dept. of Computer Science and Engineering} \\
\textit{Penn State University}\\
University Park, PA, USA \\
tlp@cse.psu.edu}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
}

\maketitle

\begin{abstract}
Peer-to-peer (p2p) networks and Mobile ad hoc networks (MANET) have been widely studied. However, a real-world deployment for the masses has remained elusive. Ever-increasing density of mobile devices, especially in urban areas, has given rise to new applications of p2p communication. However, the modern smartphone platforms have limited support for such communications. Further, the issues of battery life, range, and security remain unaddressed. A key question then is, what kinds of applications can the modern mobile platforms support and what challenges remain? This paper describes our experiences in deploying a real-world peer-to-peer delay-tolerant network to millions of users for relaying severe weather information, the challenges faced, and the approach taken.
\end{abstract}

\begin{IEEEkeywords}
peer-to-peer systems, mobile ad hoc network, delay-tolerant network
\end{IEEEkeywords}

\section{Introduction}
MANETs have been greatly studied given their decentralized nature and potential for new applications \cite{loo-manet-2011,perkins-ad-hoc-2001}. Most of the prior work has focused on valuable analytical and simulation-based study of MANET behavior \cite{zhang-topology-manet-2015,marti-misbehavior-manet-2000,mauve-pos-routing-manet-2001}. Given the outstanding practical challenges of physical nodes, real-world implementations have been limited and have not reached mass scale \cite{kiess-manet-impl-2007}. However, with the growth of smartphones, large-scale real-world implementations may become feasible. Mobile devices with programmable platforms such as android and iOS have steadily grown over the last decade, surpassing the 2 billion mark \footnote{https://www.statista.com/statistics/330695/number-of-smartphone-users-worldwide/}. This paper describes a real-world implementation of a p2p delay-tolerant network, called Mesh Network Alerts (MNA), for relaying severe weather information to millions of mobile device users as part of the Weather Channel app\footnote{https://weather.com/apps/ibm/meshnetworkalerts}. Specifically, we describe the key challenges in such an implementation, approaches taken to address them, experimental results, and a characterization of the outstanding challenges.

\section{Practical Challenges}
We focus on three main challenges and describe a system implementation for addressing them. One, even when a user is not interacting with the weather app, or worse yet, when the device is not being used at all, the devices must actively discover peer devices to send and receive potentially life-saving weather information. Although modern mobile operating systems such as Android and iOS offer APIs to discover and advertise information to peer devices over WiFi and Bluetooth interfaces, peer-to-peer connections do not work when the same APIs are accessed while the app is in the background. Prior works, widely document these challenges and take the approach of having special access on the devices, e.g., jail-breaking or rooting \cite{}. We overcame these challenges via innovative techniques, without resorting to hacks that may violate user security or Appstore guidelines. With our technique, exchange of information happens without making network connections. As we describe later, this is achieved by splitting the messages into small enough chunks so that they can be stuffed into service advertisements themselves and broadcast over multiple advertisements in a quick succession.

Two, since devices may be offline when new weather information arrives, MNA on each device must remain active at all times to be able to discover new information as soon as it arrives. However, the MNA activity must keep the device battery consumption to a minimum. This is a challenge because advertising and discovery are power-hungry operations over the radio channels. Our approach here is two-pronged. Firstly, via extensive experiments on a large number of devices, we fine-tune the algorithm governing the intervals at which discovery and advertisements occur. In a nutshell, receiving new information during a period causes MNA to be more aggressive in discovery and advertisement. Similarly, lack of new information for a period makes the device less aggressive. Secondly, we allow ``wake up'' messages to be broadcast in the network ahead of an anticipated severe weather event. When devices receive such messages, they schedule themselves to remain aggressive during the specified window of time. Outside of this window, the device can afford to have long sleep cycles and conserve power. With these techniques, our testing shows less than 1\% battery consumption per hour on most device models.

Three, given the heterogeneity of devices and operating system distributions in the market, it is quite a challenge to test whether MNA functions as expected on a given device. Further, running test scenarios on a p2p network at large-scale is non-trivial given that a large number of devices need to take specific coordinated action followed by coordinated observations to determine whether a test passes or fails. This is not a challenge in traditional mobile application development as the software is confined within a single connected device. We developed test automation tools and processes such that multiple devices can be controlled from a single test station and follow prescribed steps to generate, send, and receives messages to play out a test scenario. Finally, the framework allows automated analysis of the observations to determine the test result. This capability was instrumental in uncovering bugs at a fast pace to meet the aggressive timeline and avoid the cost of acquisition. Further, the automation framework is general and can be expressly applied to test other apps in this fashion.

Lastly, as devices with mesh advertise on a continuous basis, it may be possible for a malicious attacker to listen for such advertisements and reverse-engineer the message formats and protocols used. Then, the attackers may generate fake messages and advertise them, e.g., a fake tornado alert. Such false messages cannot be distinguished from the real ones, and MNA will end up propagating them to as many devices as possible, ``poisoning'' the network. In general, veracity of such information cannot be independently verified in open decentralized distributed systems and previous work on peer-to-peer networks do not address this challenge. In our approach, since the origin of weather information is the Weather Channel service, digital signatures can be attached to all weather alerts broadcast from the service. The mobile application is distributed with the corresponding public key so that digital signatures from the service can be verified. If a message fails such a verification, it is discarded and not forwarded any further.

\section{System Architecture}

\section{Experimental Results}

\section*{Acknowledgment}

\bibliographystyle{abbrv}
\bibliography{smartcomp19}

\end{document}

### 2019-02-09 paper.tex
Added some experimental results
Fixing author format

@@
 \title{Large-scale peer-to-peer network for mobile platforms: Challenges and Experiences\\
-\thanks{DAIS-ITA}
+%\thanks{DAIS-ITA}
 }
 
-\author{\IEEEauthorblockN{Nirmit Desai}
+\author{\IEEEauthorblockN{Nirmit Desai, Wendy Chong, Heather Achilles, Shahrokh Daijavad}
 \IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
@@
 Yorktown Heights, NY, USA \\
-nirmit.desai@us.ibm.com}
+\{nirmit.desai, wendych, hachilles, shahrokh\}@us.ibm.com}
 \and
-\IEEEauthorblockN{Wendy Chong}
-\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
+%\IEEEauthorblockN{Wendy Chong}
+%\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
 %\textit{name of organization (of Aff.)}\\
-Yorktown Heights, NY, USA \\
-wendych@us.ibm.com}
-\and
-\IEEEauthorblockN{Heather Achilles}
-\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
+%Yorktown Heights, NY, USA \\
+%wendych@us.ibm.com}
+%\and
+%\IEEEauthorblockN{Heather Achilles}
+%\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
 %\textit{name of organization (of Af}\\
-Yorktown Heights, NY, USA \\
-hachilles@us.ibm.com}
+%Yorktown Heights, NY, USA \\
+%hachilles@us.ibm.com}
+%\and
+%\IEEEauthorblockN{Shahrokh Daijavad}
+%\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
+%\textit{Penn}\\
+%Yorktown Heights, NY, USA \\
+%shahrokh@us.ibm.com}
 \and
@@
 tlp@cse.psu.edu}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+%\and
+%\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
+%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%\textit{name of organization (of Aff.)}\\
+%City, Country \\
+%email address}
+%\and
+%\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
+%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%\textit{name of organization (of Aff.)}\\
+%City, Country \\
+%email address}
 }
@@
 
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/e2e_latency}}
+\caption{End-to-end latency}
+\label{fig:e2e}
+\end{figure}
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/duplicates}}
+\caption{Duplicity due to flooding}
+\label{fig:dup}
+\end{figure}
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/hops}}
+\caption{Average hop count}
+\label{fig:hop}
+\end{figure}
+
 \section*{Acknowledgment}
-
+This research was sponsored by the U.S.
+Army Research Laboratory and the U.K.
+Ministry of Defence under Agree- ment Number W911NF-16-3-0001.
+The views and conclusions contained in this document are those of the authors and should not be interpreted as representing the official policies, either expressed or implied, of the U.S.
+Army Research Laboratory, the U.S.
+Government, the U.K.
+Ministry of Defence or the U.K.
+Government.
+The U.S.
+and U.K.
+Governments are au- thorized to reproduce and distribute reprints for Government purposes notwithstanding any copyright notation hereon.
 \bibliographystyle{abbrv}

### 2019-02-10 paper.tex
Added intro, story outline

@@
 \begin{abstract}
-Peer-to-peer (p2p) networks and Mobile ad hoc networks (MANET) have been widely studied.
-However, a real-world deployment for the masses has remained elusive.
-Ever-increasing density of mobile devices, especially in urban areas, has given rise to new applications of p2p communication.
-However, the modern smartphone platforms have limited support for such communications.
-Further, the issues of battery life, range, and security remain unaddressed.
-A key question then is, what kinds of applications can the modern mobile platforms support and what challenges remain?
-This paper describes our experiences in deploying a real-world peer-to-peer delay-tolerant network to millions of users for relaying severe weather information, the challenges faced, and the approach taken.
+Peer-to-peer (p2p) networks and Mobile ad hoc networks (MANET) have
+been widely studied. However, a real-world deployment for the masses
+has remained elusive. Ever-increasing density of mobile devices,
+especially in urban areas, has given rise to new applications of p2p
+communication. However, the modern smartphone platforms have limited
+support for such communications. Further, the issues of battery life,
+range, and security remain unaddressed. A key question then is, what
+kinds of applications can the modern mobile platforms support and what
+challenges remain? This paper identifies a class of applications and
+presents a novel p2p architecture called Mesh Network Alerts (MNA) to
+support them.  We describe our experiences in deploying MNA as a
+real-world peer-to-peer network to millions of users for relaying
+severe weather information along with the challenges faced, and the
+approaches for addressing them.
 \end{abstract}
@@
 \section{Introduction}
-MANETs have been greatly studied given their decentralized nature and potential for new applications \cite{loo-manet-2011,perkins-ad-hoc-2001}.
-Most of the prior work has focused on valuable analytical and simulation-based study of MANET behavior \cite{zhang-topology-manet-2015,marti-misbehavior-manet-2000,mauve-pos-routing-manet-2001}.
-Given the outstanding practical challenges of physical nodes, real-world implementations have been limited and have not reached mass scale \cite{kiess-manet-impl-2007}.
-However, with the growth of smartphones, large-scale real-world implementations may become feasible.
-Mobile devices with programmable platforms such as android and iOS have steadily grown over the last decade, surpassing the 2 billion mark \footnote{https://www.statista.com/statistics/330695/number-of-smartphone-users-worldwide/}.
-This paper describes a real-world implementation of a p2p delay-tolerant network, called Mesh Network Alerts (MNA), for relaying severe weather information to millions of mobile device users as part of the Weather Channel app\footnote{https://weather.com/apps/ibm/meshnetworkalerts}.
-Specifically, we describe the key challenges in such an implementation, approaches taken to address them, experimental results, and a characterization of the outstanding challenges.
+Mobile devices with programmable platforms such as android and iOS
+have steadily grown over the last decade, surpassing the 2 billion
+mark \footnote{https://www.statista.com/statistics/330695/number-of-smartphone-users-worldwide/}. MANETs
+have been greatly studied given their decentralized nature and
+potential for new applications
+\cite{loo-manet-2011,perkins-ad-hoc-2001}. Most of the prior work on
+p2p networks has focused on valuable analytical and simulation-based
+study of MANET behavior
+\cite{zhang-topology-manet-2015,marti-misbehavior-manet-2000,mauve-pos-routing-manet-2001}. Given
+the outstanding practical challenges of physical nodes, real-world
+implementations have been limited and have not reached mass scale
+\cite{kiess-manet-impl-2007}. However, with the growth of smartphones,
+large-scale real-world implementations may become feasible. This paper
+describes a real-world implementation of a p2p delay-tolerant network,
+called Mesh Network Alerts (MNA), for relaying severe weather
+information to millions of mobile device users as part of the Weather
+Channel app\footnote{https://weather.com/apps/ibm/meshnetworkalerts}
+on both android and iOS platforms.
 
+Before describing MNA, it is critical to identify applications that
+need p2p communication, given pervasive Internet connectivity. Doing
+so enables us to define key characteristics of such applications and
+focus on the challenges in meeting them.  This paper focuses on two
+separate classes of applications: communication in
+(a) disaster-affected or remote areas and (b) congested networks in
+densely populated areas, e.g., sports arenas.  The following are the
+key characteristics in these scenarios:
+
+\begin{itemize}
+\item No secondary communication infrastructure such as WiFi
+  access points to fall back on
+\item Network nodes are mobile, pattern of mobility is not
+  predictable
+\item New information may arrive at any time
+\item Trustworthy information is scarce, misinformation and
+  rumours are common place
+\item  Small payloads suffice in many cases and information
+  stays relevant for a few minutes
+\item  Device battery is a scarce resource, power supply
+  for recharging may not be available
+\item  Devices are owned by citizens, deployment of
+  special-purpose devices is cost prohibitive
+\end{itemize}
+
+The above needs are well-recognized in the industry with several
+ambitious attempts to address them, e.g., Google Loon
+project\footnote{https://loon.co} and Facebook
+Aquilla\footnote{https://en.wikipedia.org/wiki/Facebook\_Aquila},
+though with limited impact given. Leveraging user mobile devices as
+peer nodes for a large-scale deployment has been attempted before,
+e.g., the Serval project \cite{gardner-stephen-serval-2011}. Serval
+mesh enables p2p communication over on-device WiFi radio, but requires
+root access to the device via jailbreaking. Although significant
+leassons have been learned through these attempts, a mass-scale p2p
+network for such applications remains elusive.
+
+A vast majority of the literature has focused on a traditional model
+of stateful and reliable networking. Specifically, the nodes maintain
+connectivity with peers, routing is optimized with techniques based on
+link state or distance vectors
+\cite{clausen-olsr-2003,perkins-aodv-2003} focusing on optimizing the
+network utilization. Given the application characteristics above, this
+paper identifies main practical challenges associated with modern
+device platforms and finds novel ways to overcome them.  This leads to
+a new paradigm in p2p networking that is connection-less,
+delay-tolerant, and zero-routing.  With extensive experiments and
+experience of deploying to almost $10$ million users, MNA represents a
+way forward for large-scale p2p networks.
+
+In summary, this paper makes the following key contributions:
+\begin{itemize}
+\item Identification of a class of applications for p2p and their key
+  characteristics
+\item A deeper investigation of the practical challenges in supporting
+  the above class of applications
+\item A novel p2p networking implementation using multiple radio
+  channels for modern mobile platforms (Android and iOS)
+\item Experimental evaluation and deployment statistics
+\end{itemize}
+
+In the following, Section~\ref{sec:challenges} describes the practical
+challenges. Section~\ref{sec:architecture} outlines the architectural
+details of MNA along with platform-specific implementation issues for
+Android and iOS. Experimental evaluation and deployment statistics are
+presented in Section~\ref{sec:eval}.  A deeper look at the literature
+and contrast to MNA is summarized in Section~\ref{sec:related} with
+conclusions in Section~\ref{sec:conclude}.
+%
 \section{Practical Challenges}
-We focus on three main challenges and describe a system implementation for addressing them.
-One, even when a user is not interacting with the weather app, or worse yet, when the device is not being used at all, the devices must actively discover peer devices to send and receive potentially life-saving weather information.
-Although modern mobile operating systems such as Android and iOS offer APIs to discover and advertise information to peer devices over WiFi and Bluetooth interfaces, peer-to-peer connections do not work when the same APIs are accessed while the app is in the background.
-Prior works, widely document these challenges and take the approach of having special access on the devices, e.g., jail-breaking or rooting \cite{}.
-We overcame these challenges via innovative techniques, without resorting to hacks that may violate user security or Appstore guidelines.
-With our technique, exchange of information happens without making network connections.
-As we describe later, this is achieved by splitting the messages into small enough chunks so that they can be stuffed into service advertisements themselves and broadcast over multiple advertisements in a quick succession.
-
-Two, since devices may be offline when new weather information arrives, MNA on each device must remain active at all times to be able to discover new information as soon as it arrives.
-However, the MNA activity must keep the device battery consumption to a minimum.
-This is a challenge because advertising and discovery are power-hungry operations over the radio channels.
-Our approach here is two-pronged.
-Firstly, via extensive experiments on a large number of devices, we fine-tune the algorithm governing the intervals at which discovery and advertisements occur.
-In a nutshell, receiving new information during a period causes MNA to be more aggressive in discovery and advertisement.
-Similarly, lack of new information for a period makes the device less aggressive.
-Secondly, we allow ``wake up'' messages to be broadcast in the network ahead of an anticipated severe weather event.
-When devices receive such messages, they schedule themselves to remain aggressive during the specified window of time.
-Outside of this window, the device can afford to have long sleep cycles and conserve power.
-With these techniques, our testing shows less than 1\% battery consumption per hour on most device models.
-
-Three, given the heterogeneity of devices and operating system distributions in the market, it is quite a challenge to test whether MNA functions as expected on a given device.
-Further, running test scenarios on a p2p network at large-scale is non-trivial given that a large number of devices need to take specific coordinated action followed by coordinated observations to determine whether a test passes or fails.
-This is not a challenge in traditional mobile application development as the software is confined within a single connected device.
-We developed test automation tools and processes such that multiple devices can be controlled from a single test station and follow prescribed steps to generate, send, and receives messages to play out a test scenario.
-Finally, the framework allows automated analysis of the observations to determine the test result.
-This capability was instrumental in uncovering bugs at a fast pace to meet the aggressive timeline and avoid the cost of acquisition.
-Further, the automation framework is general and can be expressly applied to test other apps in this fashion.
-
-Lastly, as devices with mesh advertise on a continuous basis, it may be possible for a malicious attacker to listen for such advertisements and reverse-engineer the message formats and protocols used.
-Then, the attackers may generate fake messages and advertise them, e.g., a fake tornado alert.
-Such false messages cannot be distinguished from the real ones, and MNA will end up propagating them to as many devices as possible, ``poisoning'' the network.
-In general, veracity of such information cannot be independently verified in open decentralized distributed systems and previous work on peer-to-peer networks do not address this challenge.
-In our approach, since the origin of weather information is the Weather Channel service, digital signatures can be attached to all weather alerts broadcast from the service.
-The mobile application is distributed with the corresponding public key so that digital signatures from the service can be verified.
-If a message fails such a verification, it is discarded and not forwarded any further.
+\label{sec:challenges}
+%
+We focus on four main challenges and describe a system implementation
+for addressing them. 
+%
+\subsection{Background execution} 
+%
+Even when a user is not interacting with the weather app, or worse
+yet, when the device is not being used at all, the devices must
+actively discover peer devices to send and receive potentially
+life-saving weather information. Although modern mobile operating
+systems such as Android and iOS offer APIs to discover and advertise
+information to peer devices over WiFi and Bluetooth interfaces,
+peer-to-peer connections do not work when the same APIs are accessed
+while the app is in the background. Prior works, widely document these
+challenges and take the approach of having special access on the
+devices, e.g., jail-breaking or rooting
+\cite{gardner-stephen-serval-2011}. We overcame these challenges via
+innovative techniques, without resorting to hacks that may violate
+user security or Appstore guidelines. With our technique, exchange of
+information happens without making network connections. As we describe
+later, this is achieved by splitting the messages into small enough
+chunks so that they can be stuffed into service advertisements
+themselves and broadcast over multiple advertisements in a quick
+succession.
+%
+\subsection{Power constraints}
+%
+Since devices may be offline when new weather information
+arrives, MNA on each device must remain active at all times to be able
+to discover new information as soon as it arrives. However, the MNA
+activity must keep the device battery consumption to a minimum. This
+is a challenge because advertising and discovery are power-hungry
+operations over the radio channels. Our approach here is
+two-pronged. Firstly, via extensive experiments on a large number of
+devices, we fine-tune the algorithm governing the intervals at which
+discovery and advertisements occur. In a nutshell, receiving new
+information during a period causes MNA to be more aggressive in
+discovery and advertisement. Similarly, lack of new information for a
+period makes the device less aggressive. Secondly, we allow ``wake
+up'' messages to be broadcast in the network ahead of an anticipated
+severe weather event. When devices receive such messages, they
+schedule themselves to remain aggressive during the specified window
+of time. Outside of this window, the device can afford to have long
+sleep cycles and conserve power. With these techniques, our testing
+shows less than 1\% battery consumption per hour on most device
+models.
+%
+\subsection{Testing p2p networks}
+%
+Given the heterogeneity of devices and operating system
+distributions in the market, it is quite a challenge to test whether
+MNA functions as expected on a given device. Further, running test
+scenarios on a p2p network at large-scale is non-trivial given that a
+large number of devices need to take specific coordinated action
+followed by coordinated observations to determine whether a test
+passes or fails. This is not a challenge in traditional mobile
+application development as the software is confined within a single
+connected device. We developed test automation tools and processes
+such that multiple devices can be controlled from a single test
+station and follow prescribed steps to generate, send, and receives
+messages to play out a test scenario. Finally, the framework allows
+automated analysis of the observations to determine the test
+result. This capability was instrumental in uncovering bugs at a fast
+pace to meet the aggressive timeline and avoid the cost of
+acquisition. Further, the automation framework is general and can be
+expressly applied to test other apps in this fashion.
+%
+\subsection{Trust in information}
+%
+As devices with mesh advertise on a continuous basis, it may
+be possible for a malicious attacker to listen for such advertisements
+and reverse-engineer the message formats and protocols used. Then, the
+attackers may generate fake messages and advertise them, e.g., a fake
+tornado alert. Such false messages cannot be distinguished from the
+real ones, and MNA will end up propagating them to as many devices as
+possible, ``poisoning'' the network. In general, veracity of such
+information cannot be independently verified in open decentralized
+distributed systems and previous work on peer-to-peer networks do not
+address this challenge. In our approach, since the origin of weather
+information is the Weather Channel service, digital signatures can be
+attached to all weather alerts broadcast from the service. The mobile
+application is distributed with the corresponding public key so that
+digital signatures from the service can be verified. If a message
+fails such a verification, it is discarded and not forwarded any
+further.
 
 \section{System Architecture}
+\label{sec:architecture}
 
 \section{Experimental Results}
+\label{sec:eval}
 
@@
 
+\section{Related Work}
+\label{sec:related}
+
+\section{Conclusions and Future Work}
+\label{sec:conclude}
+
 \section*{Acknowledgment}
-This research was sponsored by the U.S.
-Army Research Laboratory and the U.K.
-Ministry of Defence under Agree- ment Number W911NF-16-3-0001.
-The views and conclusions contained in this document are those of the authors and should not be interpreted as representing the official policies, either expressed or implied, of the U.S.
-Army Research Laboratory, the U.S.
-Government, the U.K.
-Ministry of Defence or the U.K.
-Government.
-The U.S.
-and U.K.
-Governments are au- thorized to reproduce and distribute reprints for Government purposes notwithstanding any copyright notation hereon.
-\bibliographystyle{abbrv}
-\bibliography{smartcomp19}
+This research was sponsored by the U.S. Army Research Laboratory and
+the U.K. Ministry of Defence under Agree- ment Number
+W911NF-16-3-0001. The views and conclusions contained in this document
+are those of the authors and should not be interpreted as representing
+the official policies, either expressed or implied, of the U.S. Army
+Research Laboratory, the U.S. Government, the U.K. Ministry of Defence
+or the U.K. Government. The U.S. and U.K. Governments are au- thorized
+to reproduce and distribute reprints for Government purposes
+notwithstanding any copyright notation hereon.
+\bibliographystyle{abbrv} \bibliography{smartcomp19}
 

### 2019-02-10 paper.tex
Finalized section 1 and 2

@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\newcommand{\taskprefix}{CH}
+\newcommand{\tasksize}{\normalsize}
+\newenvironment{taskenv}[1]{\begin{list}{{\tasksize\sc \theenumi.}}{\usecounter{enumi}
+      \settowidth{\labelwidth}{{\tasksize\sc \taskprefix#1-99}}
+      \setlength{\leftmargin}{\labelwidth}
+      %\addtolength{\leftmargin}{2.0\labelsep}
+}}{\end{list}}
+\newcounter{task}
+\setcounter{task}{-1}
+\newcommand{\btask}{\begin{taskenv}{\taskprefix}\setcounter{enumi}{\value{task}}\renewcommand{\theenumi}{\taskprefix$_{\arabic{enumi}}$}}
+\newcommand{\etask}{\setcounter{task}{\value{enumi}}\renewcommand{\theenumi}{\arabic{enumi}.}\end{taskenv}}
+
 \begin{document}
@@
 
-\begin{itemize}
-\item No secondary communication infrastructure such as WiFi
+%
+\btask
+%
+\item\label{c:0} No secondary communication infrastructure such as WiFi
   access points to fall back on
-\item Network nodes are mobile, pattern of mobility is not
+\item\label{c:1} Network nodes are mobile, pattern of mobility is not
   predictable
-\item New information may arrive at any time
-\item Trustworthy information is scarce, misinformation and
+\item\label{c:2} New information may arrive at any time
+\item\label{c:3} Trustworthy information is scarce, misinformation and
   rumours are common place
-\item  Small payloads suffice in many cases and information
+\item\label{c:4}  Small payloads suffice in many cases and information
   stays relevant for a few minutes
-\item  Device battery is a scarce resource, power supply
+\item\label{c:5}  Device battery is a scarce resource, power supply
   for recharging may not be available
-\item  Devices are owned by citizens, deployment of
+\item\label{c:6}  Devices are owned by citizens, deployment of
   special-purpose devices is cost prohibitive
-\end{itemize}
+%
+\etask
+%
 
@@
 Aquilla\footnote{https://en.wikipedia.org/wiki/Facebook\_Aquila},
-though with limited impact given. Leveraging user mobile devices as
-peer nodes for a large-scale deployment has been attempted before,
-e.g., the Serval project \cite{gardner-stephen-serval-2011}. Serval
-mesh enables p2p communication over on-device WiFi radio, but requires
-root access to the device via jailbreaking. Although significant
-leassons have been learned through these attempts, a mass-scale p2p
-network for such applications remains elusive.
+though with limited impact. Leveraging user mobile devices as peer
+nodes for a large-scale deployment has been another theme in the prior
+works, e.g., the Serval project
+\cite{gardner-stephen-serval-2011}. Serval mesh enables p2p
+communication over on-device WiFi radio, but requires root access to
+the device via jailbreaking. Although significant leassons have been
+learned through these attempts, a mass-scale p2p network for such
+applications remains elusive.
 
@@
 device platforms and finds novel ways to overcome them.  This leads to
-a new paradigm in p2p networking that is connection-less,
-delay-tolerant, and zero-routing.  With extensive experiments and
-experience of deploying to almost $10$ million users, MNA represents a
-way forward for large-scale p2p networks.
+MNA --- a new paradigm in p2p networking that employs connection-less,
+delay-tolerant, and zero-routing overhead protocols.  MNA is
+implemented on both Android and iOS as an SDK and integrated with the
+Weather Channel mobile apps. With extensive experiments and experience
+of deploying to almost $10$ million users, MNA represents a way
+forward for large-scale p2p networks.
 
@@
   the above class of applications
-\item A novel p2p networking implementation using multiple radio
+\item A novel p2p networking SDK using multiple radio
   channels for modern mobile platforms (Android and iOS)
@@
 details of MNA along with platform-specific implementation issues for
-Android and iOS. Experimental evaluation and deployment statistics are
-presented in Section~\ref{sec:eval}.  A deeper look at the literature
-and contrast to MNA is summarized in Section~\ref{sec:related} with
-conclusions in Section~\ref{sec:conclude}.
+Android and iOS in addressing the challenges. Experimental evaluation
+and deployment statistics are presented in Section~\ref{sec:eval}.  A
+deeper look at the literature and contrast to MNA is summarized in
+Section~\ref{sec:related} with conclusions in
+Section~\ref{sec:conclude}.
 %
@@
 %
-We focus on four main challenges and describe a system implementation
-for addressing them. 
+We present the main challenges for modern mobile device platforms in
+meeting the classes of applications described above. These have been
+uncovered via extensive experiments and in some cases include direct
+feedback from the makers of Android and iOS.
 %
-\subsection{Background execution} 
+\subsection{Operating system constraints}
+\label{ch:os}
 %
-Even when a user is not interacting with the weather app, or worse
-yet, when the device is not being used at all, the devices must
-actively discover peer devices to send and receive potentially
+Due to \ref{c:2}, even when a user is not interacting with the weather
+app, or worse yet, when the device is not being used at all, the
+devices must discover peer devices to receive and formward potentially
 life-saving weather information. Although modern mobile operating
@@
 devices, e.g., jail-breaking or rooting
-\cite{gardner-stephen-serval-2011}. We overcame these challenges via
-innovative techniques, without resorting to hacks that may violate
-user security or Appstore guidelines. With our technique, exchange of
-information happens without making network connections. As we describe
-later, this is achieved by splitting the messages into small enough
-chunks so that they can be stuffed into service advertisements
-themselves and broadcast over multiple advertisements in a quick
-succession.
+\cite{gardner-stephen-serval-2011}. Clearly, such an approach does not
+scale to mass adoption.
 %
 \subsection{Power constraints}
+\label{ch:power}
 %
-Since devices may be offline when new weather information
-arrives, MNA on each device must remain active at all times to be able
-to discover new information as soon as it arrives. However, the MNA
-activity must keep the device battery consumption to a minimum. This
-is a challenge because advertising and discovery are power-hungry
-operations over the radio channels. Our approach here is
-two-pronged. Firstly, via extensive experiments on a large number of
-devices, we fine-tune the algorithm governing the intervals at which
-discovery and advertisements occur. In a nutshell, receiving new
-information during a period causes MNA to be more aggressive in
-discovery and advertisement. Similarly, lack of new information for a
-period makes the device less aggressive. Secondly, we allow ``wake
-up'' messages to be broadcast in the network ahead of an anticipated
-severe weather event. When devices receive such messages, they
-schedule themselves to remain aggressive during the specified window
-of time. Outside of this window, the device can afford to have long
-sleep cycles and conserve power. With these techniques, our testing
-shows less than 1\% battery consumption per hour on most device
-models.
+Since devices may be offline when new weather information arrives, MNA
+processes on each peer must remain active at all times to be able to
+discover new information as soon as it arrives. Further, as there is
+no back up infrastructure (\ref{c:0}) and a set of peers in range can
+change at any time (\ref{c:1}), each peer is responsible for
+constantly forwarding available information to other peers via
+advertisements. However, due to \ref{c:5}, the MNA activity must keep
+the device battery consumption to a minimum. This is a challenge
+because advertising and discovery are power-hungry operations over the
+radio channels.
 %
@@
 %
-Given the heterogeneity of devices and operating system
-distributions in the market, it is quite a challenge to test whether
-MNA functions as expected on a given device. Further, running test
+Given the heterogeneity of devices and operating system distributions
+owned by users (\ref{c:6}), it is challenging to test whether or not
+MNA functions as expected on a single device. Further, running test
 scenarios on a p2p network at large-scale is non-trivial given that a
@@
 followed by coordinated observations to determine whether a test
-passes or fails. This is not a challenge in traditional mobile
-application development as the software is confined within a single
-connected device. We developed test automation tools and processes
-such that multiple devices can be controlled from a single test
-station and follow prescribed steps to generate, send, and receives
-messages to play out a test scenario. Finally, the framework allows
-automated analysis of the observations to determine the test
-result. This capability was instrumental in uncovering bugs at a fast
-pace to meet the aggressive timeline and avoid the cost of
-acquisition. Further, the automation framework is general and can be
-expressly applied to test other apps in this fashion.
+passes or fails. Further, since range and mobility affect p2p
+communications and they are unpredictable (\ref{c:1}), it is important
+to run test cases under various mobility patterns across all nodes in
+the network. Emulated mobility frameworks such as CORE
+\cite{arenholz-core-2008} and EMANE \cite{emane} fall short as the
+connection latencies and wireless transmission are specific to device
+model and radio. This is not a challenge in traditional mobile
+application development as the application functionality is confined
+within a single connected device.
 %
@@
 %
-As devices with mesh advertise on a continuous basis, it may
-be possible for a malicious attacker to listen for such advertisements
-and reverse-engineer the message formats and protocols used. Then, the
-attackers may generate fake messages and advertise them, e.g., a fake
-tornado alert. Such false messages cannot be distinguished from the
-real ones, and MNA will end up propagating them to as many devices as
-possible, ``poisoning'' the network. In general, veracity of such
-information cannot be independently verified in open decentralized
-distributed systems and previous work on peer-to-peer networks do not
-address this challenge. In our approach, since the origin of weather
-information is the Weather Channel service, digital signatures can be
-attached to all weather alerts broadcast from the service. The mobile
-application is distributed with the corresponding public key so that
-digital signatures from the service can be verified. If a message
-fails such a verification, it is discarded and not forwarded any
-further.
-
+As user devices with MNA advertise on a continuous basis over
+well-known radio protocols, it may be possible for a malicious
+attacker to listen for such advertisements and reverse-engineer the
+message formats and protocols used. Then, the attackers may generate
+fake messages and advertise them, e.g., a fake tornado alert. Given a
+lack of trusted information in such scenarios (\ref{c:3}),
+misinformation campaigns can have disastrous consequences.  In open
+decentralized systems, such false messages cannot be distinguished
+from the real ones, and MNA will end up propagating them to as many
+devices as possible, ``poisoning'' the network. In general, veracity
+of such information cannot be independently verified in open
+decentralized distributed systems and previous work on peer-to-peer
+networks do not address this challenge.
+%
 \section{System Architecture}
 \label{sec:architecture}
+%
 
+We overcame these challenges via innovative techniques, without
+resorting to hacks that may violate user security or Appstore
+guidelines. With our technique, exchange of information happens
+without making network connections. As we describe later, this is
+achieved by splitting the messages into small enough chunks so that
+they can be stuffed into service advertisements themselves and
+broadcast over multiple advertisements in a quick succession.
+
+Our approach here is two-pronged. Firstly, via extensive experiments
+on a large number of devices, we fine-tune the algorithms governing
+the intervals at which discovery and advertisements occur. In a
+nutshell, receiving new information during a period causes MNA to be
+more aggressive in discovery and advertisement. Similarly, lack of new
+information for a period makes the device less aggressive. Secondly,
+we allow ``wake up'' messages to be broadcast in the network ahead of
+an anticipated severe weather event. When devices receive such
+messages, they schedule themselves to remain aggressive during the
+specified window of time. Outside of this window, the device can
+afford to have long sleep cycles and conserve power. With these
+techniques, our testing shows less than 1\% battery consumption per
+hour on most device models.
+
+We developed test automation tools and processes such that multiple
+devices can be controlled from a single test station and follow
+prescribed steps to generate, send, and receives messages to play out
+a test scenario. Finally, the framework allows automated analysis of
+the observations to determine the test result. This capability was
+instrumental in uncovering bugs at a fast pace to meet the aggressive
+timeline and avoid the cost of acquisition. Further, the automation
+framework is general and can be expressly applied to test other apps
+in this fashion.
+
+In our approach, since the origin of weather information is the
+Weather Channel service, digital signatures can be attached to all
+weather alerts broadcast from the service. The mobile application is
+distributed with the corresponding public key so that digital
+signatures from the service can be verified. If a message fails such a
+verification, it is discarded and not forwarded any further.
 \section{Experimental Results}

### 2019-02-10 paper.tex
Added initial related works draft

@@
 
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/test_arch}}
+\caption{Automated testing system for MNA}
+\label{fig:test_arch}
+\end{figure}
+
 We overcame these challenges via innovative techniques, without
@@
 
+%
 \section{Related Work}
 \label{sec:related}
-
+%
+Apart from the state-of-the-art cited earlier, there have been other
+notable attempts at practical large-scale deployments and we describe
+them here. Before mid-2000’s, most of the research on MANETs was based
+on Department of Defense requirements, until commodity multi-hop ad
+hoc networks began to be considered \cite{bruno-mesh-2005}. However,
+it took another decade before wireless mesh networking was used
+commercially to enable smartphones to connect via Bluetooth and WiFi
+in a popular application called FireChat \cite{firechat}. The success
+of FireChat, partially due to the news coverage of its use in
+political situations in which governments restricted access to the
+Internet, has led to many alternatives in the past few years. An
+up-to-date list of such applications is available
+here \footnote{https://alternativeto.net/software/firechat-by-open-garden/}
 \section{Conclusions and Future Work}

### 2019-02-10 paper.tex
Added all figs and experimental results

@@
 \usepackage{amsmath,amssymb,amsfonts}
+\usepackage{array,multirow,graphicx}
+\usepackage{float}
 \usepackage{algorithmic}
-\usepackage{graphicx}
 \usepackage{textcomp}
@@
 verification, it is discarded and not forwarded any further.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/android_ios}}
+\caption{Interoperability matrix between Android and iOS devices over BTLE}
+\label{fig:android_ios}
+\end{figure}
+
+\begin{table}[H]
+\caption{Android protocols and range results}
+\centering
+\begin{tabular}{|c|l|r|r|r|c|l|}
+\hline
+& \multicolumn{1}{c|}{Protocol} & \multicolumn{1}{c|}{Throughput} & \multicolumn{1}{c|}{Battery} & \multicolumn{1}{c|}{Range} & \multicolumn{1}{c|}{iOS} & \multicolumn{1}{c|}{Issues}\\
+&                               & \multicolumn{1}{c|}{Avg. kbps}  & \multicolumn{1}{c|}{per hour} & \multicolumn{1}{c|}{LoS ft} &                  &                            \\
+\hline
+\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{WiFi}}} & \textbf{DNS} & 0.2 & 2\% & 600 & No & \\
+& WiDi & 2000 & 3\% & 600 & No & Needs DNS\\
+& Hotspot & 2000 & 5\% & 600 & No & Permission\\
+\hline
+\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{BT}}} & Classic & 50 & 2\% & 600 & No & Unstable\\
+& \textbf{Nearby} & 50 & 2\% & 600 & No & \\
+& \textbf{BTLE} & 50 & 2\% & 600 & Yes & \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+\caption{iOS protocols and range results}
+\centering
+\begin{tabular}{|c|l|r|r|r|c|l|}
+\hline
+& \multicolumn{1}{c|}{Protocol} & \multicolumn{1}{c|}{Throughput} & \multicolumn{1}{c|}{Battery} & \multicolumn{1}{c|}{Range} & \multicolumn{1}{c|}{Android} & \multicolumn{1}{c|}{Issues}\\
+&                               & \multicolumn{1}{c|}{Avg. kbps}  & \multicolumn{1}{c|}{per hour} & \multicolumn{1}{c|}{LoS ft} &                  &                            \\
+\hline
+\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{WiFi}}} & Bonjour & 0.2 & 5\% & 200 & No & Battery\\
+& MPC & 2000 & 5\% & 200 & No & Battery\\
+& WiDi & 2000 & 5\% & 200 & No & Battery\\
+\hline
+& \textbf{BTLE} & 50 & 2\% & 800 & Yes & \\
+\hline
+\end{tabular}
+\end{table}
+
+%
 \section{Experimental Results}
 \label{sec:eval}
+%
+
+%
+
+%
+
+%
+\subsection{Pilot test results}
+\label{sec:romania}
+%
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/romania_latency}}
+\caption{Latency results from a 25-user pilot}
+\label{fig:romania_lat}
+\end{figure}
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/romania_activity}}
+\caption{MNA activity during the pilot}
+\label{fig:romania_act}
+\end{figure}
+
+About 13\% of the messages were received at least once (depends on
+density of users in the building). Users actively used the devices for
+~5\% of the time p2p NSD achieved lower end-to-end latencies than BT
+Classic, perhaps due to its broadcast model and wide support across
+devices. Packet sizes did not affect single-hop latencies on p2p DNS,
+perhaps because most of the messages fit within single advert. Packet
+sizes did not affect single-hop latencies on BT Classic, perhaps
+because Bluetooth has a much higher bandwidth 2 out of 25 devices have
+their clocks set wrong (confusing latency results). More than 10\% of
+the messages were received in the last 5 min of message expiry,
+implying messages were discarded due to expirty, even when they did
+not reach all devices
+
+%
+\subsection{Automated test results}
+\label{sec:automated}
+%
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/variety_e2e_latency}}
+\caption{Comparison of WiFi and Bluetooth interfaces in terms of
+  end-to-end latency}
+\label{fig:variety_e2e}
+\end{figure}
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/variety_hops}}
+\caption{Comparison of WiFi and Bluetooth interfaces in terms of
+  network topology}
+\label{fig:variety_hops}
+\end{figure}
 
@@
 \centerline{\includegraphics[width=\columnwidth]{figs/e2e_latency}}
-\caption{End-to-end latency}
+\caption{Comparison of WiFi and Bluetooth Nearby interfaces in terms
+  of end-to-end latency}
 \label{fig:e2e}

### 2019-02-11 paper.tex
Added system architecture

@@
 
-\title{Large-scale peer-to-peer network for mobile platforms: Challenges and Experiences\\
-%\thanks{DAIS-ITA}
-}
+\title{Large-scale hybrid ad hoc network for mobile platforms: Challenges and Experiences}
 
@@
 support for such communications. Further, the issues of battery life,
-range, and security remain unaddressed. A key question then is, what
+range, and trust remain unaddressed. A key question then is, what
 kinds of applications can the modern mobile platforms support and what
 challenges remain? This paper identifies a class of applications and
-presents a novel p2p architecture called Mesh Network Alerts (MNA) to
-support them.  We describe our experiences in deploying MNA as a
-real-world peer-to-peer network to millions of users for relaying
-severe weather information along with the challenges faced, and the
-approaches for addressing them.
+presents a novel center-to-peer-to-peer (c2p2p) architecture called
+Mesh Network Alerts (MNA) to support them.  We describe our
+experiences in deploying MNA as a real-world system to millions of
+users for relaying severe weather information along with the
+challenges faced, and the approaches for addressing them.
 \end{abstract}
@@
 mark \footnote{https://www.statista.com/statistics/330695/number-of-smartphone-users-worldwide/}. MANETs
-have been greatly studied given their decentralized nature and
+have been widely studied given their decentralized nature and
 potential for new applications
 \cite{loo-manet-2011,perkins-ad-hoc-2001}. Most of the prior work on
-p2p networks has focused on valuable analytical and simulation-based
-study of MANET behavior
+p2p networks has focused on analytical and simulation-based study of
+MANET behavior
 \cite{zhang-topology-manet-2015,marti-misbehavior-manet-2000,mauve-pos-routing-manet-2001}. Given
-the outstanding practical challenges of physical nodes, real-world
-implementations have been limited and have not reached mass scale
-\cite{kiess-manet-impl-2007}. However, with the growth of smartphones,
-large-scale real-world implementations may become feasible. This paper
-describes a real-world implementation of a p2p delay-tolerant network,
-called Mesh Network Alerts (MNA), for relaying severe weather
-information to millions of mobile device users as part of the Weather
-Channel app\footnote{https://weather.com/apps/ibm/meshnetworkalerts}
-on both android and iOS platforms.
+the outstanding practical challenges in deployoing a large number of
+physical nodes, real-world implementations have been limited and have
+not reached mass scale \cite{kiess-manet-impl-2007}. However, with the
+growth of smartphones, large-scale real-world implementations may
+become feasible. This paper describes a real-world implementation of a
+p2p delay-tolerant network, called Mesh Network Alerts (MNA), for
+relaying severe weather information to millions of mobile device users
+as part of the Weather Channel
+app\footnote{https://weather.com/apps/ibm/meshnetworkalerts} on both
+android and iOS platforms.
 
@@
 %
-\item\label{c:0} No secondary communication infrastructure such as WiFi
+\item\label{c:0} No communication infrastructure such as WiFi
   access points to fall back on
@@
 \item\label{c:4}  Small payloads suffice in many cases and information
-  stays relevant for a few minutes
+  retains value for a few minutes
 \item\label{c:5}  Device battery is a scarce resource, power supply
@@
 
-The above needs are well-recognized in the industry with several
-ambitious attempts to address them, e.g., Google Loon
-project\footnote{https://loon.co} and Facebook
+The above needs are well-recognized in the industry as well as
+academia with several ambitious attempts to address them, e.g., Google
+Loon project\footnote{https://loon.co} and Facebook
 Aquilla\footnote{https://en.wikipedia.org/wiki/Facebook\_Aquila},
@@
 A vast majority of the literature has focused on a traditional model
-of stateful and reliable networking. Specifically, the nodes maintain
-connectivity with peers, routing is optimized with techniques based on
-link state or distance vectors
+of stateful, fully decentralized, reliable networking. Specifically,
+the nodes maintain connectivity with peers and routing is optimized
+with techniques based on link state or distance vectors
 \cite{clausen-olsr-2003,perkins-aodv-2003} focusing on optimizing the
-network utilization. Given the application characteristics above, this
-paper identifies main practical challenges associated with modern
-device platforms and finds novel ways to overcome them.  This leads to
-MNA --- a new paradigm in p2p networking that employs connection-less,
-delay-tolerant, and zero-routing overhead protocols.  MNA is
-implemented on both Android and iOS as an SDK and integrated with the
-Weather Channel mobile apps. With extensive experiments and experience
-of deploying to almost $10$ million users, MNA represents a way
-forward for large-scale p2p networks.
+network utilization.
 
-In summary, this paper makes the following key contributions:
+Given the application characteristics above, this paper identifies
+main practical challenges associated with modern device platforms and
+finds novel ways to overcome them.  This leads to MNA --- a new
+paradigm in p2p networking that employs a hybrid, delay-tolerant, and
+zero-routing overhead architecture.  Unlike previous MANET
+architectures, MNA is a hybrid of a centralized and a decentralized
+architecture, called \emph{center-to-peer-to-peer (c2p2p)}. A central service
+is leveraged as the trusted source of information while the
+information is propagated in a decentralized fashion by peers. Such an
+architecture allows MNA to bypass the key issue of trust deficit in
+open decentralized systems and scale while retaining the ability of
+infrastructure-less communication in many application scenarios.
+
+MNA is implemented on both Android and iOS as an SDK and integrated
+with the Weather Channel mobile apps. With extensive experiments and
+experience of deploying to almost $10$ million users, MNA represents a
+way forward for large-scale p2p networks.
+
+This paper makes the following key contributions:
 \begin{itemize}
@@
   the above class of applications
-\item A novel p2p networking SDK using multiple radio
+\item A novel c2p2p architecture implementation using multiple radio
   channels for modern mobile platforms (Android and iOS)
@@
 %
-\section{Practical Challenges}
+\section{Challenges}
 \label{sec:challenges}
@@
 We present the main challenges for modern mobile device platforms in
-meeting the classes of applications described above. These have been
+supporting the classes of applications described above. These have been
 uncovered via extensive experiments and in some cases include direct
-feedback from the makers of Android and iOS.
+feedback from the developers of Android and iOS.
 %
@@
 %
-Due to \ref{c:2}, even when a user is not interacting with the weather
-app, or worse yet, when the device is not being used at all, the
-devices must discover peer devices to receive and formward potentially
-life-saving weather information. Although modern mobile operating
-systems such as Android and iOS offer APIs to discover and advertise
-information to peer devices over WiFi and Bluetooth interfaces,
-peer-to-peer connections do not work when the same APIs are accessed
-while the app is in the background. Prior works, widely document these
-challenges and take the approach of having special access on the
+Due to \ref{c:2}, even when a user is not interacting with an app, or
+worse yet, when the device is not being used at all, the devices must
+continue to discover peer devices to receive and formward
+information. Although modern mobile operating systems such as Android
+and iOS offer APIs to discover and advertise information to peer
+devices over WiFi and Bluetooth interfaces, peer-to-peer connections
+do not work reliably when the same APIs are accessed while the app is
+in the background. Further, on Android, each WiFi p2p connection must
+be explicitly approved by the device user. Prior works widely document
+these challenges and take the approach of having special access on the
 devices, e.g., jail-breaking or rooting
 \cite{gardner-stephen-serval-2011}. Clearly, such an approach does not
-scale to mass adoption.
+scale to mass adoption.\\
+%
 %
@@
 Since devices may be offline when new weather information arrives, MNA
-processes on each peer must remain active at all times to be able to
-discover new information as soon as it arrives. Further, as there is
-no back up infrastructure (\ref{c:0}) and a set of peers in range can
-change at any time (\ref{c:1}), each peer is responsible for
-constantly forwarding available information to other peers via
+on each peer must remain active at all times to be able to discover
+new information as soon as it arrives. Further, as there is no back up
+infrastructure (\ref{c:0}) and a set of peers in range can change at
+any time (\ref{c:1}), each peer is responsible for constantly
+forwarding available information to other peers via
 advertisements. However, due to \ref{c:5}, the MNA activity must keep
@@
 %
-Given the heterogeneity of devices and operating system distributions
-owned by users (\ref{c:6}), it is challenging to test whether or not
+Given the heterogeneity of devices owned by users and operating system
+distributions (\ref{c:6}), it is challenging to test whether or not
 MNA functions as expected on a single device. Further, running test
@@
 %
-As user devices with MNA advertise on a continuous basis over
-well-known radio protocols, it may be possible for a malicious
-attacker to listen for such advertisements and reverse-engineer the
-message formats and protocols used. Then, the attackers may generate
-fake messages and advertise them, e.g., a fake tornado alert. Given a
-lack of trusted information in such scenarios (\ref{c:3}),
-misinformation campaigns can have disastrous consequences.  In open
-decentralized systems, such false messages cannot be distinguished
-from the real ones, and MNA will end up propagating them to as many
-devices as possible, ``poisoning'' the network. In general, veracity
-of such information cannot be independently verified in open
-decentralized distributed systems and previous work on peer-to-peer
-networks do not address this challenge.
+As user devices with MNA advertise on unsecure wireless protocols, it
+may be possible for a malicious attacker to listen for such
+advertisements and reverse-engineer the message formats and protocols
+used. Then, the attackers may generate fake messages and advertise
+them, e.g., a fake tornado alert. Given a lack of trusted information
+in such scenarios (\ref{c:3}), misinformation campaigns can have
+disastrous consequences.  In open decentralized systems, such false
+messages cannot be distinguished from the real ones, and MNA will end
+up propagating them to as many devices as possible, ``poisoning'' the
+network. In general, veracity of such information cannot be
+independently verified in open decentralized systems and previous works
+on peer-to-peer networks do not address this challenge.
 %
-\section{System Architecture}
+\section{C2P2P System Architecture}
 \label{sec:architecture}
 %
-
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=\columnwidth]{figs/test_arch}}
-\caption{Automated testing system for MNA}
-\label{fig:test_arch}
+\centerline{\includegraphics[width=\columnwidth]{figs/arch}}
+\caption{c2p2p system architecture}
+\label{fig:arch}
 \end{figure}
 
-We overcame these challenges via innovative techniques, without
-resorting to hacks that may violate user security or Appstore
-guidelines. With our technique, exchange of information happens
-without making network connections. As we describe later, this is
-achieved by splitting the messages into small enough chunks so that
-they can be stuffed into service advertisements themselves and
-broadcast over multiple advertisements in a quick succession.
+To address (or bypass) the challenges identified above, this paper
+proposes c2p2p -- a hybrid architecture that leverages a central
+service as the sole source of trusted information as depicted in
+Figure~\ref{fig:arch}. An application-specific central service, e.g.,
+weather.com backend, originates new information as a push message and
+assigns it a unique identifier. Next, the central service produces a
+digital signature using its private key and appends it to the message
+payload. Finally, header parameters specifying TTL (duration after
+which the message expires), peer identifier of the central service,
+destination peer identifier list (for unicast) or a special identifier
+(for broadcast), and the time of origin are added to the message and
+the message is sent to a push notification service for distribution.
 
-Our approach here is two-pronged. Firstly, via extensive experiments
-on a large number of devices, we fine-tune the algorithms governing
-the intervals at which discovery and advertisements occur. In a
-nutshell, receiving new information during a period causes MNA to be
-more aggressive in discovery and advertisement. Similarly, lack of new
-information for a period makes the device less aggressive. Secondly,
-we allow ``wake up'' messages to be broadcast in the network ahead of
-an anticipated severe weather event. When devices receive such
-messages, they schedule themselves to remain aggressive during the
-specified window of time. Outside of this window, the device can
-afford to have long sleep cycles and conserve power. With these
-techniques, our testing shows less than 1\% battery consumption per
-hour on most device models.
+Mobile applications integrate MNA as an SDK including the public key
+of the corresponding central service. Peers having Internet
+connectivity receive the push messages and verify the digital
+signature to protect against spread of misinformation.  If the message
+has not expired and is destined to other peers, then the receiving
+peers start forwarding the message to other peers. Depending on the
+protocol being used, such forwarding happens as a unicast or
+broadcast.
 
+Since new information can arrive at any time or new peers may come
+into proximity, and message transmissions are not reliable, all peers
+store unexpired and verified messages and continue to forward them to
+proximal peers repeatedly. This implies that peers may receive the
+same message more than once but since the messages are globally
+unique, duplicates can be identified and ignored.  A peer sending a
+message adds its own peer identifier to a list of forwarders appended
+to the message header.  A receiving peer thus knows all the peers that
+already have the message and stops repeating the message to those
+peers, controlling the flooding.  As there are no control messages or
+any other overhead for routing, we call this \emph{zero-overhead}
+routing.
+
+The following describes how the c2p2p architecture addresses the
+challenges identified above along with known limitations. Further, to
+realize this architecture on mobile devices, we evaluated a variety of
+protocols on both Android and iOS. As described later, WiFi DNS on
+Android, Bluetooth Nearby on Android, and Blutooth low energy (BTLE)
+on both Android and iOS are the only protocols that were found
+effective. Actual techniques for discovery, advertisement, and
+connection vary across these protocols and are described next.
+%
+\subsection{Addressing the challenges}
+\label{sec:address}
+%
+\noindent\textbf{Operating system constraints} MNA addresses these
+challenges via innovative techniques, without resorting to hacks that
+may violate user security or App-store guidelines. DNS service
+discovery is a widely supported protocol for all platforms
+\cite{cheshire-dns-sd-2013}.
+
+On Android, WiFi DNS improvises on this standard such that the
+exchange of information happens without making network connections at
+all. This is achieved by splitting the message payloads into small
+enough chunks and stuffing them into service advertisements as txt
+records and broadcast over multiple advertisements in a quick
+succession. Also, we employ foreground services to keep the discovery
+and advertisements active indefinitely.
+
+On iOS, the only protocol that stays reliably active in background
+without violating iOS developer guidelines is BTLE. Unlike Android,
+iOS applications do not need to request discovery and advertisements
+contunually. Once a BTLE Peripheral (role that has information) is
+advertised and a Central (role that receives information) requests
+discovery, the app is allowed to go to sleep. When a matching central
+or a peripheral device is found, iOS wakes up the application app to
+handle such an event, even when the application is in the
+background. MNA builds on this and leverages acynchronous dispatch
+queues to turn each peer into a Central as well as a Peripheral
+simultaneously. This architecture allows bidirectional data transfers
+without concurrency issues.
+
+\noindent\textbf{Power constraints} Due to indefinite forground
+services and continuous discovery and advertisement, this is primarily
+an issue on Android. Our approach here is two-pronged. Firstly, via
+extensive experiments on a large number of devices, we fine-tune the
+algorithms governing the intervals at which discovery and
+advertisements occur. In a nutshell, receiving new information during
+a period causes MNA to be more aggressive in discovery and
+advertisement. Similarly, lack of new information for a period makes
+the device less aggressive. Secondly, we allow ``wake up'' messages to
+be broadcast in the network ahead of an anticipated severe weather
+event. When devices receive such messages, they schedule themselves to
+remain aggressive during the specified window of time. Outside of this
+window, the device can afford to have long sleep cycles and conserve
+power. With these techniques, our testing shows less than 2\% battery
+consumption per hour on most device models.
+
+\noindent\textbf{Testing p2p networks}
 We developed test automation tools and processes such that multiple
@@
 
-In our approach, since the origin of weather information is the
-Weather Channel service, digital signatures can be attached to all
-weather alerts broadcast from the service. The mobile application is
-distributed with the corresponding public key so that digital
-signatures from the service can be verified. If a message fails such a
-verification, it is discarded and not forwarded any further.
+\noindent\textbf{Trust in information} In our approach, since the
+origin of weather information is the Weather Channel service, digital
+signatures can be attached to all weather alerts broadcast from the
+service. The mobile application is distributed with the corresponding
+public key so that digital signatures from the service can be
+verified. If a message fails such a verification, it is discarded and
+not forwarded any further.
 
@@
 \end{table}
+
+%
+\subsection{Automated test framework}
+\label{sec:test}
+%
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/test_arch}}
+\caption{Automated testing system for MNA}
+\label{fig:test_arch}
+\end{figure}
 

### 2019-02-11 paper.tex
Added algorithms for WiFi DNS

@@
 \usepackage{float}
-\usepackage{algorithmic}
+%\usepackage{algorithmic}
+\usepackage[ruled,vlined]{algorithm2e}
 \usepackage{textcomp}
-\usepackage{xcolor}
+%\usepackage{xcolor}
+\usepackage[table]{xcolor}
+\usepackage{colortbl}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 
+\definecolor{Gray}{gray}{0.85}
 \newcommand{\taskprefix}{CH}
@@
 %
-\subsection{Operating system constraints}
+\subsection{Operating system restrictions}
 \label{ch:os}
@@
 scale to mass adoption.\\
-%
 %
@@
 %
-\subsection{Addressing the challenges}
-\label{sec:address}
+\subsection{Operating system restrictions}
+\label{sec:os}
+MNA works around the background activity and user permission
+restrictions via innovative techniques, without resorting to hacks
+that may violate user security or App-store guidelines.
+Tables~\ref{tab:android} and~\ref{tab:ios} summarize the various
+protocols we have experimented with using WiFi and Bluetooth
+interfaces, along with key findings as described here.  The protocols
+actually deployed are in bold.
 %
-\noindent\textbf{Operating system constraints} MNA addresses these
-challenges via innovative techniques, without resorting to hacks that
-may violate user security or App-store guidelines. DNS service
-discovery is a widely supported protocol for all platforms
-\cite{cheshire-dns-sd-2013}.
-
-On Android, WiFi DNS improvises on this standard such that the
-exchange of information happens without making network connections at
-all. This is achieved by splitting the message payloads into small
-enough chunks and stuffing them into service advertisements as txt
-records and broadcast over multiple advertisements in a quick
-succession. Also, we employ foreground services to keep the discovery
-and advertisements active indefinitely.
-
-On iOS, the only protocol that stays reliably active in background
-without violating iOS developer guidelines is BTLE. Unlike Android,
-iOS applications do not need to request discovery and advertisements
-contunually. Once a BTLE Peripheral (role that has information) is
-advertised and a Central (role that receives information) requests
-discovery, the app is allowed to go to sleep. When a matching central
-or a peripheral device is found, iOS wakes up the application app to
-handle such an event, even when the application is in the
-background. MNA builds on this and leverages acynchronous dispatch
-queues to turn each peer into a Central as well as a Peripheral
-simultaneously. This architecture allows bidirectional data transfers
-without concurrency issues.
-
-\noindent\textbf{Power constraints} Due to indefinite forground
-services and continuous discovery and advertisement, this is primarily
-an issue on Android. Our approach here is two-pronged. Firstly, via
-extensive experiments on a large number of devices, we fine-tune the
-algorithms governing the intervals at which discovery and
-advertisements occur. In a nutshell, receiving new information during
-a period causes MNA to be more aggressive in discovery and
-advertisement. Similarly, lack of new information for a period makes
-the device less aggressive. Secondly, we allow ``wake up'' messages to
-be broadcast in the network ahead of an anticipated severe weather
-event. When devices receive such messages, they schedule themselves to
-remain aggressive during the specified window of time. Outside of this
-window, the device can afford to have long sleep cycles and conserve
-power. With these techniques, our testing shows less than 2\% battery
-consumption per hour on most device models.
-
-\noindent\textbf{Testing p2p networks}
-We developed test automation tools and processes such that multiple
-devices can be controlled from a single test station and follow
-prescribed steps to generate, send, and receives messages to play out
-a test scenario. Finally, the framework allows automated analysis of
-the observations to determine the test result. This capability was
-instrumental in uncovering bugs at a fast pace to meet the aggressive
-timeline and avoid the cost of acquisition. Further, the automation
-framework is general and can be expressly applied to test other apps
-in this fashion.
-
-\noindent\textbf{Trust in information} In our approach, since the
-origin of weather information is the Weather Channel service, digital
-signatures can be attached to all weather alerts broadcast from the
-service. The mobile application is distributed with the corresponding
-public key so that digital signatures from the service can be
-verified. If a message fails such a verification, it is discarded and
-not forwarded any further.
-
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=\columnwidth]{figs/android_ios}}
-\caption{Interoperability matrix between Android and iOS devices over BTLE}
-\label{fig:android_ios}
-\end{figure}
-
+%
 \begin{table}[H]
-\caption{Android protocols and range results}
+\caption{Android protocols and key findings}
+\label{tab:android}
 \centering
@@
 \end{table}
-
+%
+%
 \begin{table}[H]
-\caption{iOS protocols and range results}
+\caption{iOS protocols and key findings}
+\label{tab:ios}
 \centering
@@
 \hline
-\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{WiFi}}} & Bonjour & 0.2 & 5\% & 200 & No & Battery\\
+\parbox[t]{2mm}{\multirow{2}{*}{\rotatebox[origin=c]{90}{WiFi}}} & Bonjour & 0.2 & 5\% & 200 & No & Battery\\
 & MPC & 2000 & 5\% & 200 & No & Battery\\
-& WiDi & 2000 & 5\% & 200 & No & Battery\\
 \hline
-& \textbf{BTLE} & 50 & 2\% & 800 & Yes & \\
+& \textbf{BTLE} & 50 & 1\% & 800 & Yes & \\
 \hline
@@
 \end{table}
+%
 
+Line of sight range in feet was measured by manually placing devices
+at increasing distances until they fail to communicate. Battery
+consumption per hour was measured as a difference between baseline
+battery consumption and when MNA was left running for 12
+hours. Reported numbers are an average percentage consumption across
+about 50 devices on each platform.
+
+BTLE is the only cross-platform protocol, although there are
+limitations with operating system versions as shown in
+Table~\ref{tab:android_ios}. Android is denoted as A, 18 and 21 are
+correspond to Android versions 4.3 and 5.0 respectively, and C and P
+denote the roles of Central and Peripheral. Key finding is that
+Android 4.3 devices acting as Peripheral cannot communicate either as
+senders or receivers.
 %
-\subsection{Automated test framework}
-\label{sec:test}
+\begin{table}[H]
+\caption{BTLE interoperability between Android and iOS devices}
+\label{tab:android_ios}
+\centering
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\hline 
+\multicolumn{1}{|c|}{Sender} & \multicolumn{6}{c|}{Receiver} \\
+\hline
+\multicolumn{1}{|c|}{} & A.18-C & A.18-P & A.21-C & A.21-P  & iOS-C & iOS-P \\
+\hline
+A.18-C & \cellcolor{Gray}         &   No       & \cellcolor{Gray}         &    Yes       &  \cellcolor{Gray}       &   Yes      \\ 
+\hline
+A.18-P &  No        & \cellcolor{Gray}         &   No       &  \cellcolor{Gray}         &   No      &  \cellcolor{Gray}       \\ 
+\hline
+A.21-C & \cellcolor{Gray}         &   No       &  \cellcolor{Gray}        &    Yes       &  \cellcolor{Gray}       &   Yes      \\ 
+\hline
+A.21-P &  Yes        & \cellcolor{Gray}         &   Yes       &  \cellcolor{Gray}         &   Yes      & \cellcolor{Gray}        \\ 
+\hline
+iOS-C & \cellcolor{Gray}         &    No      &  \cellcolor{Gray}        &     Yes       & \cellcolor{Gray}        &   Yes      \\ 
+\hline
+iOS-P &   Yes       & \cellcolor{Gray}         &    Yes      &  \cellcolor{Gray}         &   Yes      &  \cellcolor{Gray}       \\ 
+\hline
+\end{tabular}
+\end{table}
 %
+
+WiFi Direct (WiDi) is a clever combination of Hotspot with WiFi DNS
+wherein peers randomly choose to be a Hotspot or a client, discover
+WiFi credentials over WiFi DNS discovery and make WiFi connection to
+Hotspot without needing user permission.  However, both Hotspot and
+WiDi suffer from higher battery consumption than DNS and hence were
+not deployed. All Android protocols leverage foreground services to
+continue discovery and advertisement operations in the background
+indefinitely. Deployed protocols (WiFi DNS, BT Nearby, and BTLE) are
+described in detail in the following.
+
+\noindent\textbf{WiFi DNS} DNS service discovery is a widely supported
+protocol \cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises
+on this standard such that the exchange of information happens without
+making network connections at all. This is achieved by splitting the
+message payloads into small enough chunks and stuffing them into
+service advertisements as txt records and broadcast over multiple
+advertisements in a quick succession. Since there are no connections
+being made, the messages can be exchanged without prompting the user
+to grant permissions for each connection. Since no peer connections
+are needed and advertisements and discovery are broadcast, WiFi DNS is
+a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
+discovery but is power-hungry and does not interoperate with WiFi DNS
+on Android. Algorithm~\ref{algo:dns} outlines the peer process for
+WiFi DNS.
+
+\begin{algorithm}[h]
+\label{algo:dns}
+\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
+\SetAlgoLined
+\KwIn{Peer id $p_{id}$, advert/discovery intervals $T_{ad}$, $T_{disc}$}
+ $S_{name} \gets$ \textsf{'MNA'}\;
+ $\mathbb{M} \gets \varnothing$\;
+ \ForEach{timer expiry $T_{ad}$}{
+   $\mathbb{M} \gets \mathbb{M} -$ expired messages\;
+   $\mathbb{M}_{sort} \gets$ $\mathbb{M}$ sorted by least advertised first\;
+   $B \gets$ $\mathbb{M}_{sort}$ chunked by MTU size\;
+   dnsAdvertise ($S_{name}$, $B$)\;
+ }
+ \ForEach{timer expiry $T_{disc}$}{
+   dnsDiscover ($S_{name}$)\;
+ }
+ \ForEach{discovered service callback $S_{disc}$}{
+   $B \gets S_{disc}.B$\;
+   $M_{part} \gets msg(B)$\;
+   $M \gets M + M_{part}$\; 
+   \If{$M_{part}$ was the last chunk of $M$}{
+     \If{$M$ is unexpired $\wedge$ $M$ has valid signature}{
+       $\mathbb{M} \gets \mathbb{M} \cup M$\;
+     }
+   }
+ }
+ \tcc{For testing only}
+ \ForEach{app request to send $M$}{
+   \If{$M$ has not expired}{
+     $\mathbb{M} \gets \mathbb{M} \cup M$\;
+   }
+ }
+ \caption{WiFi DNS peer algorithm}
+\end{algorithm}
+
+\noindent\textbf{BTLE} The only protocol that stays reliably active in
+background without violating iOS developer guidelines is BTLE. Unlike
+Android, iOS applications do not need to request discovery and
+advertisements contunually. Once a BTLE Peripheral (role that has
+information) is advertised and a Central (role that receives
+information) requests discovery, the app is allowed to go to
+sleep. When a matching central or a peripheral device is found, iOS
+wakes up the application app to handle such an event, even when the
+application is in the background. MNA builds on this and leverages
+acynchronous dispatch queues to turn each peer into a Central as well
+as a Peripheral simultaneously. This architecture allows bidirectional
+data transfers without concurrency issues.
+%
+\subsection{Power constraints}
+\label{sec:power}
+%
+Due to indefinite forground services and continuous discovery and
+advertisement, this is primarily an issue on Android. Our approach
+here is two-pronged. Firstly, via extensive experiments on a large
+number of devices, we fine-tune the algorithms governing the intervals
+at which discovery and advertisements occur. In a nutshell, receiving
+new information during a period causes MNA to be more aggressive in
+discovery and advertisement. Similarly, lack of new information for a
+period makes the device less aggressive. Secondly, we allow ``wake
+up'' messages to be broadcast in the network ahead of an anticipated
+severe weather event. When devices receive such messages, they
+schedule themselves to remain aggressive during the specified window
+of time. Outside of this window, the device can afford to have long
+sleep cycles and conserve power. With these techniques, our testing
+shows less than 2\% battery consumption per hour on most device
+models.
+%
+\subsection{Testing p2p networks}
+\label{sec:texting}
+%
+We developed test automation tools and processes such that multiple
+devices can be controlled from a single test station and follow
+prescribed steps to generate, send, and receives messages to play out
+a test scenario. Finally, the framework allows automated analysis of
+the observations to determine the test result. This capability was
+instrumental in uncovering bugs at a fast pace to meet the aggressive
+timeline and avoid the cost of acquisition. Further, the automation
+framework is general and can be expressly applied to test other apps
+in this fashion.
 
@@
 \end{figure}
+
+%
+\subsection{Trust in information}
+\label{sec:trust}
+%
+In our approach, since the origin of weather information is the
+Weather Channel service, digital signatures can be attached to all
+weather alerts broadcast from the service. The mobile application is
+distributed with the corresponding public key so that digital
+signatures from the service can be verified. If a message fails such a
+verification, it is discarded and not forwarded any further.
 

### 2019-02-11 paper.tex
Finalized paper structure

@@
 continue discovery and advertisement operations in the background
-indefinitely. Deployed protocols (WiFi DNS, BT Nearby, and BTLE) are
-described in detail in the following.
+indefinitely. Algorithms for WiFi DNS and BTLE are described in detail
+in Section~\ref{sec:algo}.
 
-\noindent\textbf{WiFi DNS} DNS service discovery is a widely supported
-protocol \cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises
-on this standard such that the exchange of information happens without
-making network connections at all. This is achieved by splitting the
-message payloads into small enough chunks and stuffing them into
-service advertisements as txt records and broadcast over multiple
-advertisements in a quick succession. Since there are no connections
-being made, the messages can be exchanged without prompting the user
-to grant permissions for each connection. Since no peer connections
-are needed and advertisements and discovery are broadcast, WiFi DNS is
-a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
-discovery but is power-hungry and does not interoperate with WiFi DNS
-on Android. Algorithm~\ref{algo:dns} outlines the peer process for
-WiFi DNS.
-
-\begin{algorithm}[h]
-\label{algo:dns}
-\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
-\SetAlgoLined
-\KwIn{Peer id $p_{id}$, advert/discovery intervals $T_{ad}$, $T_{disc}$}
- $S_{name} \gets$ \textsf{'MNA'}\;
- $\mathbb{M} \gets \varnothing$\;
- \ForEach{timer expiry $T_{ad}$}{
-   $\mathbb{M} \gets \mathbb{M} -$ expired messages\;
-   $\mathbb{M}_{sort} \gets$ $\mathbb{M}$ sorted by least advertised first\;
-   $B \gets$ $\mathbb{M}_{sort}$ chunked by MTU size\;
-   dnsAdvertise ($S_{name}$, $B$)\;
- }
- \ForEach{timer expiry $T_{disc}$}{
-   dnsDiscover ($S_{name}$)\;
- }
- \ForEach{discovered service callback $S_{disc}$}{
-   $B \gets S_{disc}.B$\;
-   $M_{part} \gets msg(B)$\;
-   $M \gets M + M_{part}$\; 
-   \If{$M_{part}$ was the last chunk of $M$}{
-     \If{$M$ is unexpired $\wedge$ $M$ has valid signature}{
-       $\mathbb{M} \gets \mathbb{M} \cup M$\;
-     }
-   }
- }
- \tcc{For testing only}
- \ForEach{app request to send $M$}{
-   \If{$M$ has not expired}{
-     $\mathbb{M} \gets \mathbb{M} \cup M$\;
-   }
- }
- \caption{WiFi DNS peer algorithm}
-\end{algorithm}
-
-\noindent\textbf{BTLE} The only protocol that stays reliably active in
-background without violating iOS developer guidelines is BTLE. Unlike
-Android, iOS applications do not need to request discovery and
-advertisements contunually. Once a BTLE Peripheral (role that has
-information) is advertised and a Central (role that receives
-information) requests discovery, the app is allowed to go to
-sleep. When a matching central or a peripheral device is found, iOS
-wakes up the application app to handle such an event, even when the
-application is in the background. MNA builds on this and leverages
-acynchronous dispatch queues to turn each peer into a Central as well
-as a Peripheral simultaneously. This architecture allows bidirectional
-data transfers without concurrency issues.
 %
@@
 verification, it is discarded and not forwarded any further.
+%
+\section{Algorithms}
+\label{sec:algo}
+%
+\subsection{WiFi DNS}
+\label{sec:wifi}
+%
+DNS service discovery is a widely supported
+protocol \cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises
+on this standard such that the exchange of information happens without
+making network connections at all. This is achieved by splitting the
+message payloads into small enough chunks and stuffing them into
+service advertisements as txt records and broadcast over multiple
+advertisements in a quick succession. Since there are no connections
+being made, the messages can be exchanged without prompting the user
+to grant permissions for each connection. Since no peer connections
+are needed and advertisements and discovery are broadcast, WiFi DNS is
+a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
+discovery but is power-hungry and does not interoperate with WiFi DNS
+on Android. Algorithm~\ref{algo:dns} outlines the peer process for
+WiFi DNS.
 
+\begin{algorithm}[htb!]
+\label{algo:dns}
+\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
+\SetAlgoLined
+\KwIn{Peer id $p_{id}$}
+\KwIn{App id $a_{id}$, max advert/discovery intervals $max_{ad}$, $max_{disc}$}
+\KwIn{min advert/discovery intervals $min_{ad}$, $min_{disc}$}
+ $S_{name} \gets$ \textsf{'MNA'}\;
+ $\mathbb{M} \gets \varnothing$\;
+ $T_{ad} \gets min_{ad}$\;
+ $T_{disc} \gets min_{disc}$\;
+ $New \gets 0$
+ \ForEach{timer expiry $T_{ad}$}{
+   $\mathbb{M} \gets \mathbb{M} -$ expired messages\;
+   $\mathbb{M}_{sort} \gets$ $\mathbb{M}$ sorted by least advertised first\;
+   $B \gets$ $\mathbb{M}_{sort}$ chunked by MTU size\;
+   dnsAdvertise ($S_{name}$, $B$)\;
+   \If{$New = 0$}{
+     $T_{ad} \gets min(T_{ad} \times 2, max_{ad})$\;
+   }
+ }
+ \ForEach{timer expiry $T_{disc}$}{
+   dnsDiscover ($S_{name}$)\;
+   \If{$New = 0$}{
+     $T_{disc} \gets min(T_{disc} \times 2, max_{disc})$\;
+   }
+   $New \gets 0$\;
+ }
+ \ForEach{discovered service callback $S_{disc}$}{
+   $B \gets S_{disc}.B$\;
+   $M_{part} \gets msg(B)$\;
+   $M \gets M + M_{part}$\; 
+   \If{$M_{part}$ was the last chunk of $M$}{
+     \If{$M$ is not a duplicate}{
+       \If{$M$ is unexpired $\wedge$ $M$ has valid signature}{
+         $\mathbb{M} \gets \mathbb{M} \cup M$\;
+         $T_{ad} \gets max(\frac{T_{ad}}{2}, min_{ad})$\;
+         $T_{disc} \gets max(\frac{T_{disc}}{2}, min_{disc})$\;
+         $New \gets 1$\;
+         \If{$M.app = a_{id}$}{
+           Notify app with $M$\;
+         }
+       }
+     }
+   }
+ }
+ \tcc{For testing only}
+ \ForEach{app request to send $M$}{
+   \If{$M$ has not expired}{
+     $\mathbb{M} \gets \mathbb{M} \cup M$\;
+   }
+ }
+ \caption{WiFi DNS peer algorithm}
+\end{algorithm}
 %
+\subsection{BTLE}
+\label{sec:btle}
+%
+The only protocol that stays reliably active in
+background without violating iOS developer guidelines is BTLE. Unlike
+Android, iOS applications do not need to request discovery and
+advertisements contunually. Once a BTLE Peripheral (role that has
+information) is advertised and a Central (role that receives
+information) requests discovery, the app is allowed to go to
+sleep. When a matching central or a peripheral device is found, iOS
+wakes up the application app to handle such an event, even when the
+application is in the background. MNA builds on this and leverages
+acynchronous dispatch queues to turn each peer into a Central as well
+as a Peripheral simultaneously. This architecture allows bidirectional
+data transfers without concurrency issues.
+
 \section{Experimental Results}

### 2019-02-11 paper.tex
Finalized algorithms

@@
 \hline
-\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{BT}}} & Classic & 50 & 2\% & 600 & No & Unstable\\
+\parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{BT}}} & \textbf{Classic} & 50 & 2\% & 600 & No & \\
 & \textbf{Nearby} & 50 & 2\% & 600 & No & \\
@@
 %
-DNS service discovery is a widely supported
-protocol \cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises
-on this standard such that the exchange of information happens without
-making network connections at all. This is achieved by splitting the
-message payloads into small enough chunks and stuffing them into
-service advertisements as txt records and broadcast over multiple
+DNS service discovery is a widely supported protocol
+\cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises on this
+standard such that the exchange of information happens without making
+network connections at all. This is achieved by splitting the message
+payloads into small enough chunks and stuffing them into service
+advertisements as txt records and broadcast over multiple
 advertisements in a quick succession. Since there are no connections
@@
 on Android. Algorithm~\ref{algo:dns} outlines the peer process for
-WiFi DNS.
+WiFi DNS. Power conservation via time intervals is achieved by
+manupulating $T_{ad}$ and $T_{disc}$ based on traffic.
+
+Messages from multiple applications are forwarded on the same network,
+but a peer can validate signatures only for the messages targeting an
+app it hosts. Peers simply forward messages not targeting one of their
+apps. Key benefit of this design is that the density of peers is
+available to all apps, making c2p2p scenarios viable even when an
+individual app does not have a sufficient density of users.  However,
+this also allows a malicioous jammer to generate traffic for a fake
+app and force the peers to forwad it (as they do not match any of the
+apps) until the messages expire.  Although the fake information will
+never be delivered to the apps, this issue can affect MNA throughput
+for legitimate apps. A possible solution would be to maintain a
+central registry of $a_{id}$ and distribute it to all peers across
+apps.
 
@@
    $B \gets$ $\mathbb{M}_{sort}$ chunked by MTU size\;
-   dnsAdvertise ($S_{name}$, $B$)\;
+   dnsAdvertise ($S_{name}$, $p_{id}$, $B$)\;
    \If{$New = 0$}{
@@
  \ForEach{timer expiry $T_{disc}$}{
-   dnsDiscover ($S_{name}$)\;
+   dnsDiscover ($S_{name}$, $p_{id}$)\;
    \If{$New = 0$}{
@@
    $M \gets M + M_{part}$\; 
-   \If{$M_{part}$ was the last chunk of $M$}{
-     \If{$M$ is not a duplicate}{
-       \If{$M$ is unexpired $\wedge$ $M$ has valid signature}{
-         $\mathbb{M} \gets \mathbb{M} \cup M$\;
+   \If{$M_{part}$ is not a duplicate}{
         $T_{ad} \gets max(\frac{T_{ad}}{2}, min_{ad})$\;
@@
         $New \gets 1$\;
-         \If{$M.app = a_{id}$}{
+   }
+   \If{$M_{part}$ was the last chunk of $M$}{
+     \If{$M$ is unexpired}{
+         \If{$M.app = a_{id}$ $\wedge$ $M$ signature is valid}{
              Notify app with $M$\;
          }
-       }
+         $\mathbb{M} \gets \mathbb{M} \cup M$\;
      }
@@
 data transfers without concurrency issues.
+
+\begin{algorithm}[htb!]
+\label{algo:btle}
+\DontPrintSemicolon % Some LaTeX compilers require you to use \dontprintsemicolon instead
+\SetAlgoLined
+\KwIn{Peer id $p_{id}$, App id $a_{id}$, Max peripherals $max_p$}
+ $S_{name} \gets$ \textsf{'MNA'}\;
+ $\mathbb{M} \gets \varnothing$\;
+ $C \gets$ new Central() in new thread\;
+ $P \gets$ new Peripheral() in new thread\;
+ $T \gets$ new data sender thread\;
+ $C$.bleDiscovery($p_{id}$, $S_{name}$)\;
+ $P$.bleAdvert($p_{id}$, $S_{name}$)\;
+ $P_{con} \gets 0$\;
+
+ \ForEach{Peripheral $peer$ found event}{
+   \If{$P_{con} < max_p$}{
+     $C$.connect($peer$)\;
+     $P_{con} \gets P_{con} + 1$\;
+   }
+ }
+
+ \ForEach{Peripheral $peer$ connected event}{
+   \ForEach{$M$ $\in$ $\mathbb{M}$}{
+     \If{$peer \notin M$.fwdList}{
+       $T$.send($M$, $peer$)\;
+     }
+   }
+ }
+
+ \ForEach{Peripheral $peer$ disconnected event}{
+    $P_{con} \gets P_{con} - 1$\;
+ }
+
+ \ForEach{Central $peer$ connected event}{
+   \ForEach{$M$ $\in$ $\mathbb{M}$}{
+     \If{$peer \notin M$.fwdList}{
+       $T$.send($M$, $peer$)\;
+     }
+   }
+ }
+
+ \ForEach{$M$ received event from $T$.receive()}{
+   \If{$M$ is not a duplicate $\wedge$ $M$ is unexpired}{
+     \If{$M.app = a_{id}$ $\wedge$ $M$ signature is valid}{
+       Notify app with $M$\;
+     }
+     $M$.fwdList $\gets$ $M$.fwdList $\cup$ $p_{id}$\;
+     $\mathbb{M} \gets \mathbb{M} \cup M$\;     
+     $T$.send($M$, all connected peers)\;
+   }
+ }
+ \tcc{For testing only}
+ \ForEach{app request to send $M$}{
+   \If{$M$ has not expired}{
+     $M$.fwdList $\gets$ $p_{id}$ \;
+     $\mathbb{M} \gets \mathbb{M} \cup M$\;
+     $T$.send($M$, all connected peers)\;
+   }
+ }
+ \caption{BTLE peer algorithm}
+\end{algorithm}
 

### 2019-02-11 paper.tex
All done except experimental results

@@
 relaying severe weather information to millions of mobile device users
-as part of the Weather Channel
-app\footnote{https://weather.com/apps/ibm/meshnetworkalerts} on both
-android and iOS platforms.
+as part of the Weather Channel app \cite{mna} on both android and iOS
+platforms.
 
@@
 
-The following describes how the c2p2p architecture addresses the
+The following assesses the c2p2p architecture relative to the
 challenges identified above along with known limitations. Further, to
@@
 discovery and advertisement. Similarly, lack of new information for a
-period makes the device less aggressive. Secondly, we allow ``wake
-up'' messages to be broadcast in the network ahead of an anticipated
-severe weather event. When devices receive such messages, they
-schedule themselves to remain aggressive during the specified window
-of time. Outside of this window, the device can afford to have long
-sleep cycles and conserve power. With these techniques, our testing
-shows less than 2\% battery consumption per hour on most device
-models.
+period makes the device less aggressive. Secondly, we allow the
+centrasl service to broadcast``wake up'' messages ahead of an
+anticipated severe weather event.  When devices receive such messages,
+they schedule themselves to remain aggressive during the specified
+window of time.  Outside of this window, the device can afford to have
+long sleep cycles and conserve power. With these techniques, our
+testing shows less than 2\% battery consumption per hour on most
+device models.
 %
@@
 %
-We developed test automation tools and processes such that multiple
-devices can be controlled from a single test station and follow
-prescribed steps to generate, send, and receives messages to play out
-a test scenario. Finally, the framework allows automated analysis of
-the observations to determine the test result. This capability was
-instrumental in uncovering bugs at a fast pace to meet the aggressive
-timeline and avoid the cost of acquisition. Further, the automation
-framework is general and can be expressly applied to test other apps
-in this fashion.
+We developed novel test automation tools and processes that control
+multiple devices from a single test station and follow prescribed
+steps to generate, send, and receives messages to play out a test
+scenario.  The framework allows automated analysis of the observations
+to determine the test result. This capability was instrumental in
+uncovering bugs at a fast pace with tens of physical devices emloyed
+for automated testing.  Further, the automation framework is general
+and can be expressly applied to test other apps in this fashion.
+
+Figure~\ref{fig:test_arch} shows the overall test automnation
+architecture. A tester specifies high-level test cases and expected
+results in a platform and device agnostic fashion. Mobile devices run
+a UI proxy for executing platform and device-specific UI command while
+a central laptop terminal acts as the overall orchestrator and
+translates high-level test cases and distributes them to
+devices. Since the orchestrator needs a wired access to the devices
+and the devices cannot be programmed to move (unlike drones and bots),
+testing of range and mobility patterns needs to done manually, which
+is a known limitation.
+
+A key innovation here is the platform-independent language for
+specifying high-level test cases and a small set of primitive commands
+it translates to. We describe here the list of primitive commands and
+a screen recording across seven Android devices demonstrates this
+powerful capability \cite{mna-test-demo}.
+\begin{itemize}
+\item \textsf{syncAll} Wait for all devices to reach this point
+\item \textsf{object.find} Find a UI object by name or position
+\item \textsf{object.click} Click a UI object
+\item \textsf{sleep} Wait for the specified duration
+\item \textsf{setupSystemAlert} Respond to a system popup
+\item \textsf{text.set} Highlight a text input box
+\item \textsf{text.write} Write text to the text box
+\item \textsf{swipe} Swipe to the left/right or up/down
+\end{itemize}
 
@@
 %
-In our approach, since the origin of weather information is the
-Weather Channel service, digital signatures can be attached to all
-weather alerts broadcast from the service. The mobile application is
-distributed with the corresponding public key so that digital
-signatures from the service can be verified. If a message fails such a
-verification, it is discarded and not forwarded any further.
+In c2p2p, since the origin of information is always the central
+service, digital signatures can be attached to all messages
+originating from the service. The mobile application is distributed
+with the corresponding public key so that digital signatures from the
+service can be verified.  If a message fails such a verification, it
+is discarded and not forwarded any further.
+
+Peers cannot originate messages however as then a malicious peer can
+join the network, exchange keys with other peers, start generating,
+signing, and spreading fake information. This has been a longstanding
+problem for open decentralized systems. Requiring all peers to
+register their identity with a certificate authority and reporting the
+originators of fake information may discourage malicious behavior.
+Clearly, this is even harder to solve than the spread of fake
+information over the Internet because ground truth is hard to find and
+malicious peers can report genuine sources of information as well.
 %
@@
 
+Algorithm~\ref{algo:btle} outlines the peer process for BTLE. The
+biggest difference with WiFi DNS is that discovery and advertisements
+happen only once per app launch. Secondly, the entire peer process is
+event-driven, so the app can sleep when there are no events to
+process. The list of forwarders for each message $M.fwdList$ is key in
+controlling the duplicates and in turn network utilization without
+overhead messages. In practice, limiting number of Peripherals a
+Central connects to enables reliable communications, specified by
+$max_p$.
+
 \begin{algorithm}[htb!]
@@
  $\mathbb{M} \gets \varnothing$\;
- $C \gets$ new Central() in new thread\;
- $P \gets$ new Peripheral() in new thread\;
- $T \gets$ new data sender thread\;
+ $C \gets$ new Central() in new dispatch queue\;
+ $P \gets$ new Peripheral() in new dispatch queue\;
+ $T \gets$ new data sender dispatch queue\;
  $C$.bleDiscovery($p_{id}$, $S_{name}$)\;

### 2019-02-11 paper.tex
First complete draft, needs shortening and reviewing

@@
 \cite{zhang-topology-manet-2015,marti-misbehavior-manet-2000,mauve-pos-routing-manet-2001}. Given
-the outstanding practical challenges in deployoing a large number of
+the outstanding practical challenges in deploying a large number of
 physical nodes, real-world implementations have been limited and have
@@
 communication over on-device WiFi radio, but requires root access to
-the device via jailbreaking. Although significant leassons have been
+the device via jail-breaking. Although significant lessons have been
 learned through these attempts, a mass-scale p2p network for such
@@
 worse yet, when the device is not being used at all, the devices must
-continue to discover peer devices to receive and formward
+continue to discover peer devices to receive and forward
 information. Although modern mobile operating systems such as Android
@@
 \cite{arenholz-core-2008} and EMANE \cite{emane} fall short as the
-connection latencies and wireless transmission are specific to device
+connection latency and wireless transmission are specific to device
 model and radio.  This is not a challenge in traditional mobile
@@
 %
-As user devices with MNA advertise on unsecure wireless protocols, it
+As user devices with MNA advertise on unsecured wireless protocols, it
 may be possible for a malicious attacker to listen for such
@@
 which the message expires), peer identifier of the central service,
-destination peer identifier list (for unicast) or a special identifier
+destination peer identifier list (for uni-cast) or a special identifier
 (for broadcast), and the time of origin are added to the message and
@@
 peers start forwarding the message to other peers. Depending on the
-protocol being used, such forwarding happens as a unicast or
+protocol being used, such forwarding happens as a uni-cast or
 broadcast.
@@
 protocols on both Android and iOS. As described later, WiFi DNS on
-Android, Bluetooth Nearby on Android, and Blutooth low energy (BTLE)
+Android, Bluetooth Nearby on Android, and Bluetooth low energy (BTLE)
 on both Android and iOS are the only protocols that were found
@@
 %
-Due to indefinite forground services and continuous discovery and
+Due to indefinite foreground services and continuous discovery and
 advertisement, this is primarily an issue on Android. Our approach
@@
 period makes the device less aggressive. Secondly, we allow the
-centrasl service to broadcast``wake up'' messages ahead of an
+central service to broadcast``wake up'' messages ahead of an
 anticipated severe weather event.  When devices receive such messages,
@@
 to determine the test result. This capability was instrumental in
-uncovering bugs at a fast pace with tens of physical devices emloyed
+uncovering bugs at a fast pace with tens of physical devices employed
 for automated testing.  Further, the automation framework is general
@@
 
-Figure~\ref{fig:test_arch} shows the overall test automnation
+Figure~\ref{fig:test_arch} shows the overall test automation
 architecture. A tester specifies high-level test cases and expected
@@
 a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
-discovery but is power-hungry and does not interoperate with WiFi DNS
+discovery but is power-hungry and does not inter-operate with WiFi DNS
 on Android. Algorithm~\ref{algo:dns} outlines the peer process for
 WiFi DNS. Power conservation via time intervals is achieved by
-manupulating $T_{ad}$ and $T_{disc}$ based on traffic.
+manipulating $T_{ad}$ and $T_{disc}$ based on traffic.
 
@@
 individual app does not have a sufficient density of users.  However,
-this also allows a malicioous jammer to generate traffic for a fake
-app and force the peers to forwad it (as they do not match any of the
+this also allows a malicious jammer to generate traffic for a fake
+app and force the peers to forward it (as they do not match any of the
 apps) until the messages expire.  Although the fake information will
@@
 Android, iOS applications do not need to request discovery and
-advertisements contunually. Once a BTLE Peripheral (role that has
+advertisements continually. Once a BTLE Peripheral (role that has
 information) is advertised and a Central (role that receives
@@
 application is in the background. MNA builds on this and leverages
-acynchronous dispatch queues to turn each peer into a Central as well
+asynchronous dispatch queues to turn each peer into a Central as well
 as a Peripheral simultaneously. This architecture allows bidirectional
@@
 
+%
 \section{Experimental Results}
@@
 %
-
+Although MNA is deployed to about $10$ million users, collecting a
+significant amount of data runs into privacy and scale issues. Also,
+given the challenges associated with tests involving a large number of
+physical devices along with human mobility patterns, we employ a
+three-pronged approach to evaluation: (1) Limited metrics from the MNA
+user population, (2) recruit human volunteers (colleagues at IBM) in a
+pilot to have the app on their personal devices and run tests in the
+background, and (3) use the automated test environment to run
+controlled experiments. All experiments mentioning WiFi are on the
+WiFi DNS channel and BT is Bluetooth classic.
 %
-
+\subsection{MNA limited experiment}
+\label{sec:mna}
 %
-
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=\columnwidth]{figs/mna}}
+\caption{Messages received over mesh in one week}
+\label{fig:mna}
+\end{figure}
+%
+As Figure~\ref{fig:mna} shows, measures of number of messages received
+first via MNA and the number of unique users was collected during a week in
+June 2018. Almost no MNA messages were received with a roll-out to
+$10\%$ of the users.  This makes sense given that p2p needs critical
+density of users to propagate messages over large distances.  With
+$100\%$ of the users participating, 6 million+ weather alerts were
+received first via MNA by 1.9 million+ users within the collection
+period. Even when the users are connected to the Internet, they
+may receive the alert via MNA before the push notification arrives via
+the Internet. Given that such a large population of users received it
+via MNA first implies that there is value in the c2p2p architecture
+even when Internet connectivity is available.
 %
@@
 %
+About 25 colleagues participated in a pilot on a site across two
+office buildings and about a thousand people in total.  Clearly, the
+user density in this case was low. The app installed on their personal
+devices sent a broadcast message of varying sizes every five minutes
+and statistics on MNA were collected.
+
+Overall, about 13\% of the messages were received at least once
+(depends on density of users in the building) while users actively
+used the devices for ~5\% of the time.  Figure~\ref{fig:romania_lat}
+shows number of messages received and how the single-hop latency
+varies between WiFi DNS and BT Classic and across increasing packet
+sizes. Except for the 1000 byte+ packets, WiFi DNS achieved lower
+end-to-end latencies than BT Classic, perhaps due to its broadcast
+model and wide support across devices. This also explains why more
+messages were delivered over WiFi DNS than BT Classic. Packet sizes
+did not affect single-hop latencies on p2p DNS, perhaps because most
+of the messages fit within a single advertised chunk of up to 750
+bytes.
+
 \begin{figure}[htbp]
 \centerline{\includegraphics[width=\columnwidth]{figs/romania_latency}}
-\caption{Latency results from a 25-user pilot}
+\caption{Latency as a function of packet sizes}
 \label{fig:romania_lat}
@@
 
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=\columnwidth]{figs/romania_activity}}
-\caption{MNA activity during the pilot}
-\label{fig:romania_act}
-\end{figure}
-
-About 13\% of the messages were received at least once (depends on
-density of users in the building). Users actively used the devices for
-~5\% of the time p2p NSD achieved lower end-to-end latencies than BT
-Classic, perhaps due to its broadcast model and wide support across
-devices. Packet sizes did not affect single-hop latencies on p2p DNS,
-perhaps because most of the messages fit within single advert. Packet
-sizes did not affect single-hop latencies on BT Classic, perhaps
-because Bluetooth has a much higher bandwidth 2 out of 25 devices have
-their clocks set wrong (confusing latency results). More than 10\% of
-the messages were received in the last 5 min of message expiry,
-implying messages were discarded due to expirty, even when they did
-not reach all devices
+%\begin{figure}[htbp]
+%\centerline{\includegraphics[width=\columnwidth]{figs/romania_activity}}
+%\caption{MNA activity during the pilot}
+%\label{fig:romania_act}
+%\end{figure}
 
@@
 %
+We turn to the automated testing with up to 10 physical devices to
+carry out some controlled testing and a deeper understanding of how
+the various MNA protocols behave.
+%
 \begin{figure}[htbp]
 \centerline{\includegraphics[width=\columnwidth]{figs/variety_e2e_latency}}
-\caption{Comparison of WiFi and Bluetooth interfaces in terms of
-  end-to-end latency}
+\caption{End-to-end latency, WiFi DNS vs. BT Classic}
 \label{fig:variety_e2e}
 \end{figure}
+%
+Figure~\ref{fig:variety_e2e} shows end-to-end latency results in a test
+involving three Android devices, each sending a variety of message and
+payload size configurations. Latency is shown on a logarithmic scale
+y-axis. Total load on network increases from left to right on
+x-axis while the number of messages is small for the first and third
+configurations and large for the second and fourth configuration.
 
+As expected, WiFi DNS latency increases significantly with network
+load given its very low throughput. As BT Classic cannot broadcast, it
+spends significant time making connections with peers.  Hence, WiFi
+DNS beats BT Classic in the 30x67 configuration involving a large
+number of small messages. However, when the message sizes are
+increased as in 42x117 configuration, BT Classic is almost 10x faster
+than WiFi DNS due to superior
+throughput. Figure~\ref{fig:variety_hops} confirms this difference in
+network topology by measuring average hop counts. WiFi DNS always
+needs fewer hops to deliver messages than BT Classic.
+%
 \begin{figure}[htbp]
 \centerline{\includegraphics[width=\columnwidth]{figs/variety_hops}}
-\caption{Comparison of WiFi and Bluetooth interfaces in terms of
-  network topology}
+\caption{Hop counts, WiFi DNS vs. BT Classic}
 \label{fig:variety_hops}
 \end{figure}
+%
+
+Next, we compare WiFi DNS and BT Nearby on Android with BTLE on iOS in
+terms of end-to-end latency as the payload size is increased as shown
+in Figure~\ref{fig:e2e}. The results were collected over three repeats
+of the experiment involving up to 10 android devices and 6 iOS
+devices. BTLE is almost 100x faster than WiFi DNS and BT Nearby. For
+Wifi DNS, as payload size approaches the max advertisement chunk size
+of 750 bytes, the end-to-end latency rises by almost 10x (100 bytes
+versus 720 bytes). Behavior of BT Nearby taking time to form
+connections but having a good throughput is evident as the end-to-end
+latency does not change even when payload sizes go up to 10KB. BTLE on
+the other hand gets 10x slower with 10KB payloads as compared to 1KB.
 
@@
 \centerline{\includegraphics[width=\columnwidth]{figs/e2e_latency}}
-\caption{Comparison of WiFi and Bluetooth Nearby interfaces in terms
-  of end-to-end latency}
+\caption{End-to-end latency, WiFi DNS vs. BT Nearby vs. BTLE}
 \label{fig:e2e}
@@
 
+As described above, MNA employs zero-overhead routing with the goal of
+maximizing coverage and minimizing latency across peers while
+tolerating topology changes.  As a result, messages may be delivered
+multiple times to the same peer as no control messages are sent to
+keep track of the peer state.  Figure~\ref{fig:dup} shows the cost of
+such an approach in terms of average number of duplicate messages
+received by peers across WiFi DNS, BT Nearby, and BTLE (iOS).
+
+For small payload sizes, WiFi DNS ends up duplicating a large number
+of messages as the broadcasts are seen by many peers and repeated
+advertisements must go on to reach newly proximal peers. The
+duplicates drop significantly with payloads approaching the
+advertisement chunk size because the fraction of transmissions that
+are successfull also drops with large payloads (multiple chunks need
+to be delivered per message). The forward list in BT Nearby and BTLE
+help control the duplicates significantly.
+
 \begin{figure}[htbp]
 \centerline{\includegraphics[width=\columnwidth]{figs/duplicates}}
-\caption{Duplicity due to flooding}
+\caption{Duplicity due to zero-overhead routing}
 \label{fig:dup}
@@
 
+Lastly, Figure~\ref{fig:hop} shows the network toplogy underlying WiFi
+DNS, BT Nearby, and BTLE. As shown earlier in
+Figure~\ref{fig:variety_hops}, WiFi DNS delivers most of the message
+with a single hop due to a broadcast protocol whereas BT Nearby and
+BTLE need as many as two hops on average.
+%
 \begin{figure}[htbp]
@@
 \end{figure}
-
+%
 %
@@
 Apart from the state-of-the-art cited earlier, there have been other
-notable attempts at practical large-scale deployments and we describe
-them here. Before mid-2000’s, most of the research on MANETs was based
-on Department of Defense requirements, until commodity multi-hop ad
-hoc networks began to be considered \cite{bruno-mesh-2005}. However,
-it took another decade before wireless mesh networking was used
-commercially to enable smartphones to connect via Bluetooth and WiFi
-in a popular application called FireChat \cite{firechat}. The success
-of FireChat, partially due to the news coverage of its use in
-political situations in which governments restricted access to the
-Internet, has led to many alternatives in the past few years. An
-up-to-date list of such applications is available
-here \footnote{https://alternativeto.net/software/firechat-by-open-garden/}
+notable attempts at practical large-scale deployments, some of which
+will be described here. Before mid-2000s, most of the research on
+MANETs was based on Department of Defense requirements, until
+commodity multi-hop ad hoc networks began to be considered
+\cite{bruno-mesh-2005}. Around that time, academia and industry
+started publishing and implementing commodity mesh networks in
+university and industrial environments. Some notable projects from
+this era are: the community-based multi-hop wireless network by Bahl
+et al. at Microsoft Research \cite{microsoft-mesh}, the Broadband and
+Wireless Network (BWN) project at Georgia Tech \cite{gatech-mesh}, the
+testbed created at UCLA Network Research Lab for Wireless Mesh
+Networks and the evaluation of its performance of real-time traffic
+\cite{chavoutier-2007}, the wireless mesh network testbed at Shanghai
+Jiao Tong University \cite{wu-mesh-2010}, and a survey paper about
+Wireless Mesh Networks and its challenges at the transport layer
+\cite{saha-mesh-2014}.
+
+Despite these efforts, it took until 2014 before wireless mesh
+networking was used commercially to enable smartphones to connect via
+Bluetooth and WiFi in a popular application called FireChat
+\cite{firechat}. The success of FireChat, partially due to the news
+coverage of its use in political situations in which governments
+restricted access to the Internet, has led to many alternatives in the
+past few years. An up-to-date list of such applications is available
+here4. With the recent introduction of p2p support at the Operating
+System level by Apple and Google for smartphones, we believe there
+will be a significant increase in both academic and industrial
+implementations of large scale mesh networks in the upcoming years.
+%
 \section{Conclusions and Future Work}
 \label{sec:conclude}
+%
+We reported on MNA, a large-scale real-world hybrid ad hoc network for
+Android and iOS platforms. Leveraging a central service as the single
+source of trusted information with a novel c2p2p architecture enables
+a class of applications to take advantage of p2p communications over
+an increasing density of mobile devices. Operating system restrictions
+are addressed or bypassed with simple and creative algorithms without
+violating user privacy and security or app deverloper
+guidelines. Extensive experiments with real users, pilot users, and
+automation show that MNA enables low throughput connectivity on modern
+devices with low battery consumption on both Android and iOS
+platforms.
 
+Although MNA represents a significant step forward in the right
+direction, there are significant limitations to be overcome. First,
+many applications require peers to originate information, e.g., asking
+for help during a natural disaster. Hence, the issue of trust in open
+decentralized systems needs to be investigated.  Second, although MNA
+delivers almost all messages sent to all peers, it cannot guarantee a
+100\% delivery rate.  In many applications, such guarantees are
+needed. These are challenging problems for further research.
+
+%
 \section*{Acknowledgment}
 This research was sponsored by the U.S. Army Research Laboratory and
-the U.K. Ministry of Defence under Agree- ment Number
+the U.K. Ministry of Defence under Agreement Number
 W911NF-16-3-0001. The views and conclusions contained in this document
@@
 Research Laboratory, the U.S. Government, the U.K. Ministry of Defence
-or the U.K. Government. The U.S. and U.K. Governments are au- thorized
+or the U.K. Government. The U.S. and U.K. Governments are Authorized
 to reproduce and distribute reprints for Government purposes
@@
 \end{document}
+
+% LocalWords:  hoc Nirmit Desai Chong Shahrokh Daijavad nirmit desai wendych BT
+% LocalWords:  hachilles shahrokh Porta MNA iOS MANETs Aquilla Serval stateful
+% LocalWords:  SDK APIs EMANE backend TTL DNS BTLE kbps LoS WiDi Hotspot MPC UI
+% LocalWords:  Bonjour interoperability orchestrator syncAll setupSystemAlert
+% LocalWords:  txt jammer MTU dnsAdvertise dnsDiscover msg fwdList bleDiscovery
+% LocalWords:  bleAdvert NSD latencies FireChat thorized smartcomp

### 2019-02-19 paper.tex
Final submitted version

@@
 \section{Introduction}
-Mobile devices with programmable platforms such as android and iOS
+Mobile devices with programmable platforms such as Android and iOS
 have steadily grown over the last decade, surpassing the 2 billion
-mark \footnote{https://www.statista.com/statistics/330695/number-of-smartphone-users-worldwide/}. MANETs
+mark\footnote{https://www.statista.com/statistics/330695/number-of-smartphone-users-worldwide/}. MANETs
 have been widely studied given their decentralized nature and
@@
 relaying severe weather information to millions of mobile device users
-as part of the Weather Channel app \cite{mna} on both android and iOS
+as part of the Weather Channel app \cite{mna} on both Android and iOS
 platforms.
@@
 Before describing MNA, it is critical to identify applications that
-need p2p communication, given pervasive Internet connectivity. Doing
-so enables us to define key characteristics of such applications and
-focus on the challenges in meeting them.  This paper focuses on two
-separate classes of applications: communication in
-(a) disaster-affected or remote areas and (b) congested networks in
+need p2p communication, given pervasive Internet connectivity and the
+ease of using it. Doing so enables us to define key characteristics of
+such applications and focus on the challenges in meeting them.  This
+paper focuses on two separate classes of applications: communication
+in (a) disaster-affected or remote areas and (b) congested networks in
 densely populated areas, e.g., sports arenas.  The following are the
@@
   access points to fall back on
-\item\label{c:1} Network nodes are mobile, pattern of mobility is not
+\item\label{c:1} Device users are mobile, pattern of mobility is not
   predictable
@@
 the nodes maintain connectivity with peers and routing is optimized
-with techniques based on link state or distance vectors
+with techniques based on link state (OLSR) or distance vectors (AODV)
 \cite{clausen-olsr-2003,perkins-aodv-2003} focusing on optimizing the
@@
 Given the application characteristics above, this paper identifies
-main practical challenges associated with modern device platforms and
-finds novel ways to overcome them.  This leads to MNA --- a new
-paradigm in p2p networking that employs a hybrid, delay-tolerant, and
-zero-routing overhead architecture.  Unlike previous MANET
-architectures, MNA is a hybrid of a centralized and a decentralized
-architecture, called \emph{center-to-peer-to-peer (c2p2p)}. A central service
-is leveraged as the trusted source of information while the
-information is propagated in a decentralized fashion by peers. Such an
-architecture allows MNA to bypass the key issue of trust deficit in
-open decentralized systems and scale while retaining the ability of
+practical challenges associated with modern device platforms and finds
+novel ways to overcome them.  This leads to MNA -- a new paradigm in
+p2p networking that employs a delay-tolerant, and zero-routing
+overhead architecture.  Unlike previous works, MNA is a hybrid of a
+centralized and a decentralized architecture, called
+\emph{center-to-peer-to-peer (c2p2p)}. A central service is leveraged
+as the trusted source of information while the information is
+propagated in a decentralized fashion by peers. Such an architecture
+allows MNA to bypass the key issue of trust deficit in open
+decentralized systems while retaining the ability of
 infrastructure-less communication in many application scenarios.
@@
   the above class of applications
-\item A novel c2p2p architecture implementation using multiple radio
-  channels for modern mobile platforms (Android and iOS)
-\item Experimental evaluation and deployment statistics
+\item A novel c2p2p implementation using multiple radio
+  channels on modern mobile platforms (Android and iOS)
+\item Experimental evaluation and real-world deployment statistics
 \end{itemize}
@@
 details of MNA along with platform-specific implementation issues for
-Android and iOS in addressing the challenges. Experimental evaluation
-and deployment statistics are presented in Section~\ref{sec:eval}.  A
-deeper look at the literature and contrast to MNA is summarized in
-Section~\ref{sec:related} with conclusions in
+Android and iOS in addressing the challenges. Section~\ref{sec:algo}
+dives into the algorithms of two of the MNA protocols. Experimental
+evaluation and deployment statistics are presented in
+Section~\ref{sec:eval}.  A deeper look at the literature and contrast
+to MNA is summarized in Section~\ref{sec:related} with conclusions in
 Section~\ref{sec:conclude}.
@@
 worse yet, when the device is not being used at all, the devices must
-continue to discover peer devices to receive and forward
+continue to discover peers to receive and forward
 information. Although modern mobile operating systems such as Android
-and iOS offer APIs to discover and advertise information to peer
-devices over WiFi and Bluetooth interfaces, peer-to-peer connections
-do not work reliably when the same APIs are accessed while the app is
-in the background. Further, on Android, each WiFi p2p connection must
-be explicitly approved by the device user. Prior works widely document
+and iOS offer APIs to discover and advertise information to peers over
+WiFi and Bluetooth interfaces, peer-to-peer connections do not work
+reliably when the same APIs are accessed while the app is in the
+background. Further, on Android, each WiFi p2p connection must be
+explicitly approved by the user. Prior works widely document
 these challenges and take the approach of having special access on the
@@
 distributions (\ref{c:6}), it is challenging to test whether or not
-MNA functions as expected on a single device. Further, running test
+MNA works as expected on a single device. Further, running test
 scenarios on a p2p network at large-scale is non-trivial given that a
-large number of devices need to take specific coordinated action
+large number of devices need to take coordinated action
 followed by coordinated observations to determine whether a test
@@
 connection latency and wireless transmission are specific to device
-model and radio.  This is not a challenge in traditional mobile
+hardware.  This is not a challenge in traditional mobile
 application development as the application functionality is confined
@@
 %
-As user devices with MNA advertise on unsecured wireless protocols, it
-may be possible for a malicious attacker to listen for such
-advertisements and reverse-engineer the message formats and protocols
-used. Then, the attackers may generate fake messages and advertise
-them, e.g., a fake tornado alert. Given a lack of trusted information
-in such scenarios (\ref{c:3}), misinformation campaigns can have
-disastrous consequences.  In open decentralized systems, such false
-messages cannot be distinguished from the real ones, and MNA will end
-up propagating them to as many devices as possible, ``poisoning'' the
+As user devices advertise on unsecured wireless protocols, it may be
+possible for a malicious attacker to listen for such advertisements
+and reverse-engineer the protocols used. Then, the attackers may
+generate fake messages and advertise them, e.g., a fake tornado
+alert. Given a lack of trusted information in such scenarios
+(\ref{c:3}), misinformation campaigns can have disastrous
+consequences.  In open decentralized systems, such false messages
+cannot be distinguished from the real ones, and MNA will end up
+propagating them to as many devices as possible, ``poisoning'' the
 network. In general, veracity of such information cannot be
-independently verified in open decentralized systems and previous works
-on peer-to-peer networks do not address this challenge.
+independently verified in open decentralized systems and previous
+works on peer-to-peer networks do not address this challenge.
 %
@@
 Figure~\ref{fig:arch}. An application-specific central service, e.g.,
-weather.com backend, originates new information as a push message and
+weather.com backend, originates new information and
 assigns it a unique identifier. Next, the central service produces a
@@
 
-Mobile applications integrate MNA as an SDK including the public key
-of the corresponding central service. Peers having Internet
-connectivity receive the push messages and verify the digital
-signature to protect against spread of misinformation.  If the message
-has not expired and is destined to other peers, then the receiving
-peers start forwarding the message to other peers. Depending on the
-protocol being used, such forwarding happens as a uni-cast or
-broadcast.
+Mobile applications integrate MNA as an SDK with the public key of the
+corresponding central service. Peers having Internet connectivity
+receive the push messages and verify the digital signature.  If the
+signature is valid, the message has not expired, and is destined to
+other peers, then the receiving peers forward the message to other
+peers. Depending on the protocol being used, such forwarding happens
+as a uni-cast or broadcast. Peers receiving the forwarded messages
+also verify the digital signature before forwarding further.
 
-Since new information can arrive at any time or new peers may come
-into proximity, and message transmissions are not reliable, all peers
-store unexpired and verified messages and continue to forward them to
-proximal peers repeatedly. This implies that peers may receive the
-same message more than once but since the messages are globally
-unique, duplicates can be identified and ignored.  A peer sending a
-message adds its own peer identifier to a list of forwarders appended
-to the message header.  A receiving peer thus knows all the peers that
-already have the message and stops repeating the message to those
-peers, controlling the flooding.  As there are no control messages or
-any other overhead for routing, we call this \emph{zero-overhead}
-routing.
+Since new information can arrive at any time, new peers may come into
+proximity, or message transmissions fail, all peers store unexpired
+and verified messages and forward them to proximal peers
+repeatedly. This implies that peers may receive a message more than
+once but since the messages are globally unique, duplicates can be
+identified and ignored.  A peer sending a message adds its own peer
+identifier to a list of forwarders appended to the message header.  A
+receiving peer thus knows all the peers that already have the message
+and stops repeating the message to those peers, controlling the
+flooding.  As there are no control messages or any other overhead for
+routing, we call this \emph{zero-overhead} routing.
 
@@
 challenges identified above along with known limitations. Further, to
-realize this architecture on mobile devices, we evaluated a variety of
+realize this architecture on mobile devices, we evaluate a variety of
 protocols on both Android and iOS. As described later, WiFi DNS on
@@
 \label{sec:os}
-MNA works around the background activity and user permission
-restrictions via innovative techniques, without resorting to hacks
-that may violate user security or App-store guidelines.
-Tables~\ref{tab:android} and~\ref{tab:ios} summarize the various
-protocols we have experimented with using WiFi and Bluetooth
-interfaces, along with key findings as described here.  The protocols
-actually deployed are in bold.
 %
@@
 %
+
+MNA works around the background activity and user permission
+restrictions via innovative techniques, without resorting to hacks
+that may violate user security or developer guidelines.
+Tables~\ref{tab:android} and~\ref{tab:ios} summarize the various
+protocols we have experimented with using WiFi and Bluetooth
+interfaces, along with key findings as described here.  The protocols
+actually deployed are in bold.
+
 %
@@
 
-Line of sight range in feet was measured by manually placing devices
-at increasing distances until they fail to communicate. Battery
-consumption per hour was measured as a difference between baseline
-battery consumption and when MNA was left running for 12
-hours. Reported numbers are an average percentage consumption across
-about 50 devices on each platform.
+Throughput is measured over a single p2p hop. Line of Sight (LoS)
+range in feet was measured by manually placing devices at increasing
+distances until they fail to communicate. Battery consumption per hour
+was measured as a difference between baseline battery consumption and
+when MNA was left running for 12 hours. Reported numbers are an
+average percentage consumption across about 50 devices on each
+platform.
 
@@
 limitations with operating system versions as shown in
-Table~\ref{tab:android_ios}. Android is denoted as A, 18 and 21 are
-correspond to Android versions 4.3 and 5.0 respectively, and C and P
-denote the roles of Central and Peripheral. Key finding is that
-Android 4.3 devices acting as Peripheral cannot communicate either as
-senders or receivers.
+Table~\ref{tab:android_ios}. Android is denoted as A, 18 and 21 API
+levels corresponding to Android versions 4.3 and 5.0 respectively, and
+C and P denote the roles of Central and Peripheral. Key finding is
+that Android 4.3 devices acting as Peripheral cannot communicate
+either as senders or receivers.
 %
@@
 they schedule themselves to remain aggressive during the specified
-window of time.  Outside of this window, the device can afford to have
+window of time.  Outside of this window, the devices can afford to have
 long sleep cycles and conserve power. With these techniques, our
@@
 
-Figure~\ref{fig:test_arch} shows the overall test automation
-architecture. A tester specifies high-level test cases and expected
-results in a platform and device agnostic fashion. Mobile devices run
-a UI proxy for executing platform and device-specific UI command while
-a central laptop terminal acts as the overall orchestrator and
-translates high-level test cases and distributes them to
-devices. Since the orchestrator needs a wired access to the devices
-and the devices cannot be programmed to move (unlike drones and bots),
-testing of range and mobility patterns needs to done manually, which
-is a known limitation.
+Figure~\ref{fig:test_arch} shows the test automation architecture. A
+tester specifies high-level test cases and expected results in a
+platform and device-agnostic fashion. Mobile devices run a UI proxy
+app for executing platform and device-specific UI commands while a
+central laptop terminal acts as the overall orchestrator. The
+orchestrator translates high-level test cases and distributes them to
+devices. Since the orchestrator needs wired access to the devices and
+the devices cannot be programmed to move, testing of range and
+mobility patterns needs to done manually, which is a known limitation.
 
@@
 it translates to. We describe here the list of primitive commands and
-a screen recording across seven Android devices demonstrates this
-powerful capability \cite{mna-test-demo}.
+have made available a screen recording across seven Android devices
+demonstrates this powerful capability \cite{mna-test-demo}.
 \begin{itemize}
-\item \textsf{syncAll} Wait for all devices to reach this point
-\item \textsf{object.find} Find a UI object by name or position
-\item \textsf{object.click} Click a UI object
-\item \textsf{sleep} Wait for the specified duration
-\item \textsf{setupSystemAlert} Respond to a system popup
-\item \textsf{text.set} Highlight a text input box
-\item \textsf{text.write} Write text to the text box
-\item \textsf{swipe} Swipe to the left/right or up/down
+\item \textsf{syncAll} -- Wait for all devices to reach this point
+\item \textsf{object.find} -- Find a UI object by name or position
+\item \textsf{object.click} -- Click a UI object
+\item \textsf{sleep} -- Wait for the specified duration
+\item \textsf{setupSystemAlert} -- Respond to a system popup
+\item \textsf{text.set} -- Highlight a text input box
+\item \textsf{text.write} -- Write text to the text box
+\item \textsf{swipe} -- Swipe to the left/right or up/down
 \end{itemize}
@@
 %
-In c2p2p, since the origin of information is always the central
-service, digital signatures can be attached to all messages
-originating from the service. The mobile application is distributed
+In c2p2p, since the central service originates all messages, it can
+digitally sign them. The mobile application is distributed
 with the corresponding public key so that digital signatures from the
@@
 information over the Internet because ground truth is hard to find and
-malicious peers can report genuine sources of information as well.
+malicious peers can report genuine sources of information as fake.
 %
@@
 \cite{cheshire-dns-sd-2013}. On Android, WiFi DNS improvises on this
-standard such that the exchange of information happens without making
-network connections at all. This is achieved by splitting the message
-payloads into small enough chunks and stuffing them into service
-advertisements as txt records and broadcast over multiple
-advertisements in a quick succession. Since there are no connections
-being made, the messages can be exchanged without prompting the user
-to grant permissions for each connection. Since no peer connections
-are needed and advertisements and discovery are broadcast, WiFi DNS is
-a broadcast protocol. On iOS, Bonjour protocol is built on DNS service
-discovery but is power-hungry and does not inter-operate with WiFi DNS
-on Android. Algorithm~\ref{algo:dns} outlines the peer process for
-WiFi DNS. Power conservation via time intervals is achieved by
-manipulating $T_{ad}$ and $T_{disc}$ based on traffic.
+standard such that communication happens without making network
+connections. This is achieved by splitting the message payloads into
+small enough chunks, stuffing them into service advertisements as txt
+records, and broadcasting them over multiple advertisements. Since
+there are no connections being made, the messages can be exchanged
+without prompting the user to grant permissions for each
+connection. Since no peer connections are needed and advertisements
+and discovery are broadcast, WiFi DNS is a broadcast
+protocol. Although the standard states that as much as 65KB of txt
+records can be added to advertisements, in practice most devices
+support only up to 750 bytes. On iOS, Bonjour protocol is built on DNS
+service discovery but is power-hungry and does not inter-operate with
+WiFi DNS on Android. Algorithm~\ref{algo:dns} outlines the peer
+process for WiFi DNS. Power conservation via time intervals is
+achieved by manipulating timers $T_{ad}$ and $T_{disc}$ based on
+traffic. $\mathbb{M}$ is the set of all unexpired messages.
 
@@
 but a peer can validate signatures only for the messages targeting an
-app it hosts. Peers simply forward messages not targeting one of their
-apps. Key benefit of this design is that the density of peers is
-available to all apps, making c2p2p scenarios viable even when an
-individual app does not have a sufficient density of users.  However,
-this also allows a malicious jammer to generate traffic for a fake
-app and force the peers to forward it (as they do not match any of the
-apps) until the messages expire.  Although the fake information will
-never be delivered to the apps, this issue can affect MNA throughput
-for legitimate apps. A possible solution would be to maintain a
-central registry of $a_{id}$ and distribute it to all peers across
-apps.
+app it hosts. This is because peers do not possess the public keys
+corresponding to the central services of other apps. Hence, peers
+simply forward messages not targeting one of their apps. Key benefit
+of this design is that the entire network of peers is available to all
+apps, making c2p2p scenarios viable even when an individual app does
+not have a sufficient density of users.  However, this also allows a
+malicious jammer to generate traffic for a fake app, forcing the peers
+to forward it until the messages expire.  Although the fake
+information will never be delivered to the apps, this affects
+throughput. A possible solution would be to distribute a registry of
+valid $a_{id}$ to all peers across apps so that they can ignore
+messages with fake $a_{id}$ and not forward them.
 
@@
 \SetAlgoLined
-\KwIn{Peer id $p_{id}$}
-\KwIn{App id $a_{id}$, max advert/discovery intervals $max_{ad}$, $max_{disc}$}
+\KwIn{Peer id $p_{id}$, App id $a_{id}$}
+\KwIn{max advert/discovery intervals $max_{ad}$, $max_{disc}$}
 \KwIn{min advert/discovery intervals $min_{ad}$, $min_{disc}$}
@@
  $T_{disc} \gets min_{disc}$\;
- $New \gets 0$
+ $New \gets 0$\;
  \ForEach{timer expiry $T_{ad}$}{
@@
 %
-The only protocol that stays reliably active in
-background without violating iOS developer guidelines is BTLE. Unlike
-Android, iOS applications do not need to request discovery and
-advertisements continually. Once a BTLE Peripheral (role that has
-information) is advertised and a Central (role that receives
-information) requests discovery, the app is allowed to go to
-sleep. When a matching central or a peripheral device is found, iOS
-wakes up the application app to handle such an event, even when the
-application is in the background. MNA builds on this and leverages
+
+The only protocol that stays reliably active in background without
+violating developer guidelines is BTLE. Unlike Android, iOS
+applications do not need to discover and advertise continually. Once a
+BTLE Peripheral (role that has information) is advertised and a
+Central (role that receives information) requests discovery, the app
+is allowed to go to sleep. When a matching Central or a Peripheral is
+found, iOS wakes up the application to handle such an event, even when
+the application is in the background. MNA builds on this and leverages
 asynchronous dispatch queues to turn each peer into a Central as well
@@
 process. The list of forwarders for each message $M.fwdList$ is key in
-controlling the duplicates and in turn network utilization without
-overhead messages. In practice, limiting number of Peripherals a
-Central connects to enables reliable communications, specified by
-$max_p$.
+controlling duplicate transmissions without overhead messages. In
+practice, limiting number of Peripherals a Central connects to enables
+reliable communications, specified by $max_p$.
 
@@
 %
-Although MNA is deployed to about $10$ million users, collecting a
-significant amount of data runs into privacy and scale issues. Also,
-given the challenges associated with tests involving a large number of
-physical devices along with human mobility patterns, we employ a
-three-pronged approach to evaluation: (1) Limited metrics from the MNA
-user population, (2) recruit human volunteers (colleagues at IBM) in a
-pilot to have the app on their personal devices and run tests in the
-background, and (3) use the automated test environment to run
-controlled experiments. All experiments mentioning WiFi are on the
-WiFi DNS channel and BT is Bluetooth classic.
+Although MNA is deployed to about $10$ million users, collecting test
+data runs into privacy and scale issues. Also, given that testing with
+a large number of physical devices along with mobility is challenging,
+we employ a three-pronged approach to evaluation: (1) gather limited
+metrics from the MNA user population, (2) recruit human volunteers in
+a pilot to have the tests run on their personal devices, and (3) use
+the automated test environment to run controlled experiments. All
+experiments mentioning WiFi are on the WiFi DNS channel and BT is
+Bluetooth classic.
 %
@@
 %
-As Figure~\ref{fig:mna} shows, measures of number of messages received
-first via MNA and the number of unique users was collected during a week in
-June 2018. Almost no MNA messages were received with a roll-out to
-$10\%$ of the users.  This makes sense given that p2p needs critical
-density of users to propagate messages over large distances.  With
-$100\%$ of the users participating, 6 million+ weather alerts were
-received first via MNA by 1.9 million+ users within the collection
-period. Even when the users are connected to the Internet, they
-may receive the alert via MNA before the push notification arrives via
-the Internet. Given that such a large population of users received it
-via MNA first implies that there is value in the c2p2p architecture
-even when Internet connectivity is available.
+As Figure~\ref{fig:mna} shows, test data comprising number of messages
+received first via MNA and the number of unique users was collected
+during a week in June 2018. Almost no MNA messages were received with
+a roll-out to $10\%$ of the users.  This makes sense given that p2p
+needs critical density of users to propagate messages over large
+distances.  With $100\%$ of the users participating, 6 million+
+weather alerts were received first via MNA by 1.9 million+ users. Even
+when the users are connected to the Internet, they may receive the
+alert via MNA before the push notification arrives via the
+Internet. Given that such a large population of users received it via
+MNA first implies that there is value in the c2p2p architecture even
+when Internet connectivity is available.
 %
@@
 user density in this case was low. The app installed on their personal
-devices sent a broadcast message of varying sizes every five minutes
+devices broadcast messages of varying sizes every five minutes
 and statistics on MNA were collected.
 
-Overall, about 13\% of the messages were received at least once
-(depends on density of users in the building) while users actively
-used the devices for ~5\% of the time.  Figure~\ref{fig:romania_lat}
-shows number of messages received and how the single-hop latency
-varies between WiFi DNS and BT Classic and across increasing packet
-sizes. Except for the 1000 byte+ packets, WiFi DNS achieved lower
-end-to-end latencies than BT Classic, perhaps due to its broadcast
-model and wide support across devices. This also explains why more
-messages were delivered over WiFi DNS than BT Classic. Packet sizes
-did not affect single-hop latencies on p2p DNS, perhaps because most
-of the messages fit within a single advertised chunk of up to 750
-bytes.
+Overall, about 13\% of the messages were received at least once while
+users actively used the devices for ~5\% of the time.
+Figure~\ref{fig:romania_lat} shows how number of messages received and
+the single-hop latency vary between WiFi DNS and BT Classic with
+increasing packet sizes. Except for the 1000 byte+ packets, WiFi DNS
+achieved lower end-to-end latencies than BT Classic, perhaps due to
+its broadcast model. This also explains why more messages were
+delivered over WiFi DNS than BT Classic. Packet sizes did not affect
+single-hop latencies on WiFi DNS, perhaps because most of the messages
+fit within a single advertised chunk.
 
@@
 We turn to the automated testing with up to 10 physical devices to
-carry out some controlled testing and a deeper understanding of how
-the various MNA protocols behave.
+carry out some controlled experiments and a deeper understanding of
+how the various MNA protocols behave. In these experiments, all
+messages has a TTL of 15 minutes and all messages were received within
+that duration except for the 10KB messages on WiFi DNS.
 %
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=\columnwidth]{figs/variety_e2e_latency}}
+\centerline{\includegraphics[width=0.8\columnwidth]{figs/variety_e2e_latency}}
 \caption{End-to-end latency, WiFi DNS vs. BT Classic}
@@
 %
-Figure~\ref{fig:variety_e2e} shows end-to-end latency results in a test
-involving three Android devices, each sending a variety of message and
-payload size configurations. Latency is shown on a logarithmic scale
-y-axis. Total load on network increases from left to right on
-x-axis while the number of messages is small for the first and third
-configurations and large for the second and fourth configuration.
+
+Figure~\ref{fig:variety_e2e} shows end-to-end latency results with
+three Android devices, each sending a varying number of messages with
+varying sizes. Latency is shown on y-axis on a logarithmic
+scale. Total load on network increases from left to right on x-axis
+while the number of messages is small for the first and third
+cases it is large for the second and fourth cases.
 
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=\columnwidth]{figs/variety_hops}}
+\centerline{\includegraphics[width=0.8\columnwidth]{figs/variety_hops}}
 \caption{Hop counts, WiFi DNS vs. BT Classic}
@@
 Next, we compare WiFi DNS and BT Nearby on Android with BTLE on iOS in
-terms of end-to-end latency as the payload size is increased as shown
+terms of end-to-end latency as the payload size increases as shown
 in Figure~\ref{fig:e2e}. The results were collected over three repeats
@@
 devices. BTLE is almost 100x faster than WiFi DNS and BT Nearby. For
-Wifi DNS, as payload size approaches the max advertisement chunk size
+WiFi DNS, as payload size approaches the max advertisement chunk limit
 of 750 bytes, the end-to-end latency rises by almost 10x (100 bytes
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=\columnwidth]{figs/e2e_latency}}
+\centerline{\includegraphics[width=0.8\columnwidth]{figs/e2e_latency}}
 \caption{End-to-end latency, WiFi DNS vs. BT Nearby vs. BTLE}
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=\columnwidth]{figs/duplicates}}
+\centerline{\includegraphics[width=0.8\columnwidth]{figs/duplicates}}
 \caption{Duplicity due to zero-overhead routing}
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=\columnwidth]{figs/hops}}
-\caption{Average hop count}
+\centerline{\includegraphics[width=0.8\columnwidth]{figs/hops}}
+\caption{Avg hop count, WiFi DNS vs. BT Nearby vs. BTLE}
 \label{fig:hop}
@@
 past few years. An up-to-date list of such applications is available
-here4. With the recent introduction of p2p support at the Operating
-System level by Apple and Google for smartphones, we believe there
-will be a significant increase in both academic and industrial
+here\footnote{https://alternativeto.net/software/firechat-by-open-garden/}.
+All of these works are fully decentralized and cannot guarantee
+veracity of information being propagated.
+
+Unlike the above-cited works, MNA employs a hybrid architecture
+focused on trusted information and low throughput low-power
+protocols. With the recent introduction of p2p support at the
+operating system level by Apple and Google for smartphones, we believe
+there will be a significant increase in both academic and industrial
 implementations of large scale mesh networks in the upcoming years.
@@
 We reported on MNA, a large-scale real-world hybrid ad hoc network for
-Android and iOS platforms. Leveraging a central service as the single
-source of trusted information with a novel c2p2p architecture enables
-a class of applications to take advantage of p2p communications over
-an increasing density of mobile devices. Operating system restrictions
-are addressed or bypassed with simple and creative algorithms without
-violating user privacy and security or app deverloper
-guidelines. Extensive experiments with real users, pilot users, and
-automation show that MNA enables low throughput connectivity on modern
-devices with low battery consumption on both Android and iOS
-platforms.
+Android and iOS platforms. Leveraging a central service as the source
+of trusted information with a novel c2p2p architecture enables a class
+of applications over an increasing density of mobile devices. MNA
+addresses or bypasses operating system restrictions with simple and
+creative algorithms without violating user privacy, security, or
+developer guidelines. Extensive experiments with real users, pilot
+users, and automation show that MNA enables low-throughput
+connectivity on modern devices with low battery consumption on both
+Android and iOS platforms.
 
